---
title: "dam_synthesis"
author: "L. Naslund"
date: "2024-08-05"
output: html_document
---

```{r set up}
library(tidyverse)
library(cowplot)

veazie <- read.csv("../2-output-data/veazie_summary_20240725.csv") %>% mutate(dam = "Veazie")
glines <- read.csv("../2-output-data/glines_summary_20240725.csv") %>% mutate(dam = "Glines Canyon")

all_res <- veazie %>% bind_rows(glines)   %>% 
  mutate(
    mean_area_mg = mean_area * 1e-6,
    ci_lwr_area_mg = ci_lwr_area * 1e-6,
    ci_upr_area_mg = ci_upr_area * 1e-6
  )

gen_theme <- function(g){
  return(g +  theme_bw()+
  theme(axis.text.x = element_text(size = 16, color = "black"), axis.text.y = element_text(size = 16, color = "black"), axis.title.x = element_blank(), axis.title.y = element_text(size =20), legend.title = element_blank(), legend.text = element_text(size = 16)))
}

before_after_df <- all_res %>% filter(
  period %in% c("before", "build"),
  variation %in% c(NA, "with degassing", "100", "three meters", "three meters angio")
) %>%
  mutate(
    flux = case_when(
      flux == "reservoir emissions" ~ "emissions",
      flux == "reservoir burial" ~ "burial",
      flux == "river CO2 emissions" ~ "river CO2",
      flux == "river CH4 emissions" ~ "river CH4",
      flux == "NEP" ~ "NEP",
      flux == "soil CH4 emissions" ~ "soil CH4",
      flux == "tree CH4 emissions" ~ "tree CH4"
    )
  ) %>%
  mutate(flux_fct = as.factor(flux) %>% fct_relevel(
    c(
      "emissions",
      "burial",
      "river CO2",
      "river CH4",
      "NEP",
      "soil CH4",
      "tree CH4"
    )
  )) %>%
  mutate(
    period_fct = case_when(period == "before" ~ "before",
                           period == "build" ~ "after") %>% fct_relevel("before", "after")
  ) 
```

```{r calculate burp scenarios}
veazie_scenarios <- all_res %>% 
  filter(dam == "Veazie", period == "burp", flux_unit == "CO2-eq")%>% 
  mutate(flux_var = paste(flux, variation)) %>% 
  filter(flux_var %in% c("exposed sediment CO2 0.083333333", "exposed sediment CH4 no_pore_0.0833333333333333", "drawdown ebullition 0", 
                         "exposed sediment CO2 5", "exposed sediment CH4 pore_5", "drawdown ebullition 10",  
                         "exposed sediment CO2 30", "exposed sediment CH4 pore_30", "drawdown ebullition 100", "eroded sediment emissions NA"))

veazie_scenarios_rw <- veazie_scenarios %>% slice(10)

veazie_scenarios <- veazie_scenarios %>% bind_rows(veazie_scenarios_rw) %>% bind_rows(veazie_scenarios_rw) %>% 
  mutate(scenario = rep(c("least", "moderate", "most"), 4))

veazie_scenarios_summary <- veazie_scenarios %>% group_by(scenario) %>% summarize(net_mg = sum(mean_area_mg), net_mg_lwr = sum(ci_lwr_area_mg), net_mg_upr = sum(ci_upr_area_mg))

glines_scenarios <- all_res %>% 
  filter(dam == "Glines Canyon", period == "burp", flux_unit == "CO2-eq")%>% 
  mutate(flux_var = paste(flux, variation)) %>% 
  filter(flux_var %in% c("exposed sediment CO2 0.0833333333333333", "exposed sediment CH4 no_pore_0.0833333333333333", "drawdown ebullition 0", 
                         "exposed sediment CO2 5", "exposed sediment CH4 pore_5", "drawdown ebullition 10",  
                         "exposed sediment CO2 30", "exposed sediment CH4 pore_30", "drawdown ebullition 100", "eroded sediment emissions NA"))

glines_scenarios_rw <- glines_scenarios %>% slice(10)

glines_scenarios <- glines_scenarios %>% bind_rows(glines_scenarios_rw) %>% bind_rows(glines_scenarios_rw) %>% 
  mutate(scenario = rep(c("least", "moderate", "most"), 4))

glines_scenarios_summary <- glines_scenarios %>% group_by(scenario) %>% summarize(net_mg = sum(mean_area_mg), net_mg_lwr = sum(ci_lwr_area_mg), net_mg_upr = sum(ci_upr_area_mg))
```


# Abstract
```{r before after df}
before_after_df %>% 
  filter(flux_unit == "CO2-eq") %>% 
  group_by(dam, period) %>% 
  summarize(mean_mg = sum(mean_area_mg), lwr_mg = sum(ci_lwr_area_mg), upr_mg = sum(ci_upr_area_mg))
```

# Results

Reservoir emissions before removal
```{r res emissions}
all_res %>% 
  filter(flux == "reservoir emissions", flux_unit == "CO2-eq") %>% 
  dplyr::select(flux, variation, dam, mean_area_mg, ci_lwr_area_mg, ci_upr_area_mg)
```

Reservoir burial before removal
```{r res burial}
all_res %>% 
  filter(flux == "reservoir burial", flux_unit == "CO2-eq") %>% 
  dplyr::select(flux, variation, dam, mean_area_mg, ci_lwr_area_mg, ci_upr_area_mg)
```

Burp summary CO2-eq
```{r burp totals}
# summary of burp emissions by scenario
veazie_scenarios_summary

# equivalent years of preremoval emissions
veazie_pre_removal_emissions <- before_after_df %>% filter(dam == "Veazie", flux_unit == "CO2-eq", flux == "emissions") %>% pull(mean_area_mg)

veazie_scenarios_summary %>% dplyr::select(scenario, net_mg) %>% mutate(yrs = net_mg/veazie_pre_removal_emissions)

# summary of burp emissions by scenario
glines_scenarios_summary

# equivalent years of preremoval emissions
glines_pre_removal_emissions <- before_after_df %>% filter(dam == "Glines Canyon", flux_unit == "CO2-eq", flux == "emissions") %>% pull(mean_area_mg)

glines_scenarios_summary %>% dplyr::select(scenario, net_mg) %>% mutate(yrs = net_mg/glines_pre_removal_emissions)
```

River CO2 emissions after removal
```{r}
all_res %>% 
  filter(flux == "river CO2 emissions", flux_unit == "CO2-eq") %>% 
  dplyr::select(flux, variation, dam, mean_area_mg, ci_lwr_area_mg, ci_upr_area_mg)
```

Terrestrial NEP after removal
```{r}
all_res %>% 
  filter(flux == "NEP", flux_unit == "CO2-eq", variation == 100) %>% 
  dplyr::select(flux, variation, dam, mean_area_mg, ci_lwr_area_mg, ci_upr_area_mg)
```

# Main text figures

Figure 2
```{r fig 2}
veazie_before.plt <- gen_theme(before_after_df %>% filter(dam == "Veazie", period == "before", flux_unit == "CO2-eq") %>% 
  ggplot(aes(flux_fct, mean_area_mg))+
  geom_hline(yintercept = 0, linetype = "dotted", color = "red", linewidth = 1)+
  geom_point(aes(), size = 2)+
  geom_errorbar(aes(x = flux_fct, ymin = ci_lwr_area_mg, ymax = ci_upr_area_mg), linewidth = 1, width = 0.5)+
  ylim(-2000, 25000)+
  ggtitle("Veazie Before")+
  ylab(expression(Mg~CO[2-eq]~yr^-1)))+
  theme(plot.title = element_text(size=20, hjust = 0.5))

veazie_after.plt <- gen_theme(before_after_df %>% filter(dam == "Veazie", period == "build", flux_unit == "CO2-eq") %>% 
  ggplot(aes(flux_fct, mean_area_mg))+
  geom_hline(yintercept = 0, linetype = "dotted", color = "red", linewidth = 1)+
  geom_point(aes(), size = 2)+
  geom_errorbar(aes(x = flux_fct, ymin = ci_lwr_area_mg, ymax = ci_upr_area_mg), linewidth = 1, width = 0.5)+
  ylim(-2000, 25000)+
  ggtitle("Veazie After")+
    scale_x_discrete(labels = c(expression(river~CO[2]), expression(river~CH[4]), "NEP", expression(soil~CH[4]), expression(tree~CH[4])))+
  ylab(expression(Mg~CO[2-eq]~yr^-1)))+
  theme(plot.title = element_text(size=20, hjust = 0.5))

glines_before.plt <- gen_theme(before_after_df %>% filter(dam == "Glines Canyon", period == "before", flux_unit == "CO2-eq") %>% 
  ggplot(aes(flux_fct, mean_area_mg))+
  geom_hline(yintercept = 0, linetype = "dotted", color = "red", linewidth = 1)+
  geom_point(aes(), size = 2)+
  geom_errorbar(aes(x = flux_fct, ymin = ci_lwr_area_mg, ymax = ci_upr_area_mg), linewidth = 1, width = 0.5)+
  ylim(-35000, 2500)+
  ggtitle("Glines Canyon Before")+
  ylab(expression(Mg~CO[2-eq]~yr^-1)))+
  theme(plot.title = element_text(size=20, hjust = 0.5))


glines_after.plt <- gen_theme(before_after_df %>% filter(dam == "Glines Canyon", period == "build", flux_unit == "CO2-eq") %>% 
  ggplot(aes(flux_fct, mean_area_mg))+
  geom_hline(yintercept = 0, linetype = "dotted", color = "red", linewidth = 1)+
  geom_point(aes(), size = 2)+
  geom_errorbar(aes(x = flux_fct, ymin = ci_lwr_area_mg, ymax = ci_upr_area_mg), linewidth = 1, width = 0.5)+
  ylim(-35000, 2500)+
  ggtitle("Glines Canyon After")+
  ylab(expression(Mg~CO[2-eq]~yr^-1)))+
  scale_x_discrete(labels = c(expression(river~CO[2]), expression(river~CH[4]), "NEP", expression(soil~CH[4]), expression(tree~CH[4])))+
    theme(plot.title = element_text(size=20, hjust = 0.5))

png(paste0("../2-output-data/before_after_flux_paper_", Sys.Date(), ".png"), width = 11, height = 9, units = "in", res = 300)
plot_grid(veazie_before.plt, veazie_after.plt, glines_before.plt, glines_after.plt, ncol = 2, rel_widths = c(0.4, 0.6))
dev.off()
```

Figure 3
```{r}
veazie_scenarios_fig <- gen_theme(ggplot(veazie_scenarios, aes(flux, mean_area_mg, color = scenario))+
  geom_point(size = 2, position = position_dodge(width = 0.7))+
  geom_errorbar(aes(ymin = ci_lwr_area_mg, ymax = ci_upr_area_mg), size = 1, position = position_dodge(width = 0.7), width = 0.5) + ylab(expression(Mg~CO[2-eq]))+ scale_color_manual(name = "scenario", values = c( "#155AAE",  "grey40", "#C34237"), labels = c("least", "moderate", "most"))+
    scale_x_discrete(labels = c(expression(atop("drawdown", "ebullition")), expression(atop("eroded sediment", "emissions")), expression(atop("exposed",sediment~CH[4])), expression(atop("exposed",sediment~CO[2]))))) + theme(legend.title = element_text(size = 16))

veazie_scenarios_summary_fig <- gen_theme(ggplot(veazie_scenarios_summary, aes(scenario, net_mg, color = scenario))+
  geom_point(size = 2, position = position_dodge(width = 0.7))+
    scale_x_discrete(labels = c("least", "moderate", "most"))+
  geom_errorbar(aes(ymin = net_mg_lwr, ymax = net_mg_upr), size = 1, position = position_dodge(width = 0.7), width = 0.5) + ylab(expression(Mg~CO[2-eq]))+ scale_color_manual(name = "scenario", values = c( "#155AAE",  "grey40", "#C34237"), labels = c("least", "moderate", "most")))+ theme(legend.position="none")

glines_scenarios_fig <- gen_theme(ggplot(glines_scenarios, aes(flux, mean_area_mg, color = scenario))+
  geom_point(size = 2, position = position_dodge(width = 0.7))+
  geom_errorbar(aes(ymin = ci_lwr_area_mg, ymax = ci_upr_area_mg), size = 1, position = position_dodge(width = 0.7), width = 0.5) + ylab(expression(Mg~CO[2-eq]))+ scale_color_manual(name = "scenario", values = c( "#155AAE",  "grey40", "#C34237"), labels = c("least", "moderate", "most"))+
    scale_x_discrete(labels = c(expression(atop("drawdown", "ebullition")), expression(atop("eroded sediment", "emissions")), expression(atop("exposed",sediment~CH[4])), expression(atop("exposed",sediment~CO[2]))))) + theme(legend.title = element_text(size = 16))

glines_scenarios_summary_fig <- gen_theme(ggplot(glines_scenarios_summary, aes(scenario, net_mg, color = scenario))+
  geom_point(size = 2, position = position_dodge(width = 0.7))+
    scale_x_discrete(labels = c("least", "moderate", "most"))+
  geom_errorbar(aes(ymin = net_mg_lwr, ymax = net_mg_upr), size = 1, position = position_dodge(width = 0.7), width = 0.5) + ylab(expression(Mg~CO[2-eq]))+ scale_color_manual(name = "scenario", values = c( "#155AAE",  "grey40", "#C34237"), labels = c("least", "moderate", "most")))+ theme(legend.position="none")

options(scipen =3)
png(paste0("../2-output-data/burp_scenarios_", Sys.Date(), ".png"), width = 14, height = 9, units = "in", res = 300)
plot_grid(veazie_scenarios_fig, veazie_scenarios_summary_fig, glines_scenarios_fig, glines_scenarios_summary_fig, rel_widths = c(0.7, 0.3), ncol = 2, labels = "AUTO", label_size = 20)
dev.off()
```

Figure 4
```{r}
before_after_fig <- gen_theme(before_after_df %>% 
  filter(flux_unit == "CO2-eq") %>% 
  group_by(dam, period) %>% 
  summarize(mean_mg = sum(mean_area_mg), lwr_mg = sum(ci_lwr_area_mg), upr_mg = sum(ci_upr_area_mg)) %>% 
  ggplot(aes(period, mean_mg))+
  geom_point(size = 2)+
  geom_hline(yintercept = 0, linetype = "dotted", color = "red", linewidth = 1)+
  geom_errorbar(aes(x = period, ymin = lwr_mg, ymax = upr_mg), linewidth = 1)+
  ylab(expression(Mg~CO[2-eq]~yr^-1))+
  facet_wrap(~dam))

png(paste0("../2-output-data/before_after_", Sys.Date(), ".png"), width = 8, height = 5, units = "in", res = 300)
before_after_fig + theme(strip.text = element_text(size = 16))
dev.off()
```

Figure 5
```{r glines trajectory}
# calculate the time to total possible carbon storage
glines_total_volume <- 52994001 # total volume (m3) from the 1921 line in Figure 10 Bountry et al. 2011 

glines_bulk_density <- 2.03e13/1.51e7 #g m-3

glines_c_content <- (320 * 1e9)/2.03e13 # Gg C to g / g sediment, Stratton et al. 2019 Table 3

glines_total_c <- (glines_total_volume * glines_bulk_density * glines_c_content) 

glines_c_rate <- glines %>% filter(flux_unit == "C", flux == "reservoir burial") %>% pull(mean_area)

# time to max C storage
glines_max_c_yr <- round(glines_total_c/glines_c_rate, digits = 0) * -1

# accumulated C before removal
glines_total_c_actual <- glines %>% filter(flux == "reservoir burial", flux_unit == "C") %>% pull(mean_area)*87 

# c storage lost in eroded C
glines_eroded_mass <- 9.1 * 1e6 * 1e6 * glines_c_content # table 4 Randle et al. 2015

# after removal storage
glines_nep <- glines %>% filter(flux == "NEP", variation == 100, flux_unit == "CO2-eq") %>% pull(mean_area) 

glines_soil_ch4 <- glines %>% filter(flux == "soil CH4 emissions", flux_unit == "CO2-eq") %>% pull(mean_area) 

# burp
glines_exp_ch4 <- glines %>% filter(flux == "exposed sediment CH4", variation == "pore_5", flux_unit == "CO2-eq") %>% pull(mean_area) 

glines_exp_co2 <-  glines %>% filter(flux == "exposed sediment CO2", variation == "5", flux_unit == "CO2-eq") %>% pull(mean_area) 

glines_drawdown <- glines %>% filter(flux == "drawdown ebullition", variation == "10", flux_unit == "CO2-eq") %>% pull(mean_area)

glines_eroded <- glines %>% filter(flux == "eroded sediment emissions", flux_unit == "CO2-eq") %>% pull(mean_area)

# after removal emissions
glines_tree_ch4 <- glines %>% filter(flux == "tree CH4 emissions", variation == "three meters", flux_unit == "CO2-eq") %>% pull(mean_area)

glines_river_co2 <- glines %>% filter(flux == "river CO2 emissions", flux_unit == "CO2-eq") %>% pull(mean_area)

glines_river_ch4 <- glines %>% filter(flux == "river CH4 emissions", flux_unit == "CO2-eq") %>% pull(mean_area)


# reservoir emissions
glines_emissions <- glines %>% filter(flux == "reservoir emissions", variation == "with degassing", flux_unit == "CO2-eq") %>% pull(mean_area)

glines_pathway_before <- data.frame(yr = seq(0, 87, by = 1)) %>% mutate(storage_rem = yr * glines_c_rate *44/12, 
                                                                        storage_stay = yr * glines_c_rate *44/12, 
                                                                        emissions_rem = yr * glines_emissions, 
                                                                        emissions_stay = yr * glines_emissions)

glines_pathway_after <- data.frame(yr = seq(87, 375, by = 1)) %>% mutate(storage_rem = ((glines_total_c_actual + glines_eroded_mass)* (44/12)) + ((glines_nep + glines_soil_ch4) * (yr-87)), 
                                                                         storage_stay = if_else(yr < glines_max_c_yr, yr * glines_c_rate *44/12, -glines_total_c * (44/12)), 
                                                                         emissions_rem = (glines_emissions * 87) + (glines_exp_ch4 + glines_exp_co2 + glines_drawdown + glines_eroded) + ((glines_tree_ch4 + glines_river_ch4 + glines_river_co2) * (yr - 87)), 
                                                                         emissions_stay = yr * glines_emissions)

glines_pathways.plt <- glines_pathway_before %>% bind_rows(glines_pathway_after) %>% 
  mutate(storage_rem = storage_rem * 1e-9, 
         storage_stay = storage_stay * 1e-9, 
         emissions_rem = emissions_rem * 1e-9, 
         emissions_stay = emissions_stay * 1e-9, 
         total_rem = storage_rem + emissions_rem, 
         total_stay = storage_stay + emissions_stay) %>% 
  ggplot()+
  #geom_hline(aes(yintercept = 0), color = "red", linetype = "dotted", linewidth = 1)+
  geom_line(aes(yr, emissions_stay), linetype = "solid", color = "grey70", linewidth = 1)+
  geom_line(aes(yr, storage_stay), linetype = "solid", color = "grey70", linewidth = 1)+
  geom_line(aes(yr, total_stay), linetype = "dashed", color = "grey70", linewidth = 1)+
  geom_line(aes(yr, storage_rem), linetype = "solid", color = "black", linewidth = 1)+
  geom_line(aes(yr, total_rem), linetype = "dashed", color = "black", linewidth = 1)+
  geom_line(aes(yr, emissions_rem), linetype = "solid", color = "black", linewidth = 1)+
  theme_bw()+
  ggtitle("Glines Canyon Dam")+
  labs(x = "Dam age (yr)", y = expression(Carbon~balance~"(Gg"~CO[2]-eq~")"))+
  theme(axis.text = element_text(size = 14, color = "black"), axis.title = element_text(size =16), plot.title = element_text(size=20, hjust = 0.5))
```

```{r veazie trajectory}
# calculate the time to total carbon storage
veazie_sed_fig5 <- read.csv("../2-output-data/veazie_sed_fig5.csv")

# total volume: couldn't find storage capacity listed anywhere so had to approximate with dam height * reservoir area
veazie_total_volume <- veazie_sed_fig5$area * 10

veazie_total_c <- veazie_total_volume * veazie_sed_fig5$dbd * veazie_sed_fig5$oc 

# years until max storage
veazie_max_c_yr <- veazie_total_c/(veazie %>% filter(flux == "reservoir burial", flux_unit == "C") %>% pull(mean_area))*-1

# actual C burial rate
veazie_c_rate <- veazie %>% filter(flux_unit == "C", flux == "reservoir burial") %>% pull(mean_area)

# accumulated C before removal
veazie_total_c_actual <- veazie %>% filter(flux == "reservoir burial", flux_unit == "C") %>% pull(mean_area)*100 

# c storage lost in eroded C
# chose the middle of the uniform distribution fit for reservoir sediment volume
veazie_eroded_mass <- 7500 * veazie_sed_fig5$dbd * veazie_sed_fig5$oc

# after removal storage
veazie_nep <- veazie %>% filter(flux == "NEP", variation == 100, flux_unit == "CO2-eq") %>% pull(mean_area) 

veazie_soil_ch4 <- veazie %>% filter(flux == "soil CH4 emissions", flux_unit == "CO2-eq") %>% pull(mean_area) 

# burp
veazie_exp_ch4 <- veazie %>% filter(flux == "exposed sediment CH4", variation == "pore_5", flux_unit == "CO2-eq") %>% pull(mean_area) 

veazie_exp_co2 <-  veazie %>% filter(flux == "exposed sediment CO2", variation == "5", flux_unit == "CO2-eq") %>% pull(mean_area) 

veazie_drawdown <- veazie %>% filter(flux == "drawdown ebullition", variation == "10", flux_unit == "CO2-eq") %>% pull(mean_area)

veazie_eroded <- veazie %>% filter(flux == "eroded sediment emissions", flux_unit == "CO2-eq") %>% pull(mean_area)

# after removal emissions
veazie_tree_ch4 <- veazie %>% filter(flux == "tree CH4 emissions", variation == "three meters angio", flux_unit == "CO2-eq") %>% pull(mean_area)

veazie_river_co2 <- veazie %>% filter(flux == "river CO2 emissions", flux_unit == "CO2-eq") %>% pull(mean_area)

veazie_river_ch4 <- veazie %>% filter(flux == "river CH4 emissions", flux_unit == "CO2-eq") %>% pull(mean_area)

# reservoir emissions
veazie_emissions <- veazie %>% filter(flux == "reservoir emissions", flux_unit == "CO2-eq") %>% pull(mean_area)

veazie_pathway_before <- data.frame(yr = seq(0, 100, by = 1)) %>% mutate(storage_rem = yr * veazie_c_rate *44/12, 
                                                                        storage_stay = yr * veazie_c_rate *44/12, 
                                                                        emissions_rem = yr * veazie_emissions, 
                                                                        emissions_stay = yr * veazie_emissions)

veazie_pathway_after <- data.frame(yr = seq(100, 375, by = 1)) %>% mutate(storage_rem = ((veazie_total_c_actual + veazie_eroded_mass)* (44/12)) + ((veazie_nep + veazie_soil_ch4) * (yr-100)), 
                                                                         storage_stay = if_else(yr < veazie_max_c_yr, yr * veazie_c_rate *44/12, -veazie_total_c * (44/12)), 
                                                                         emissions_rem = (veazie_emissions * 100) + (veazie_exp_ch4 + veazie_exp_co2 + veazie_drawdown + veazie_eroded) + ((veazie_tree_ch4 + veazie_river_ch4 + veazie_river_co2) * (yr - 100)), 
                                                                         emissions_stay = yr * veazie_emissions)

veazie_pathways.plt <- veazie_pathway_before %>% bind_rows(veazie_pathway_after) %>% 
  mutate(storage_rem = storage_rem * 1e-9, 
         storage_stay = storage_stay * 1e-9, 
         emissions_rem = emissions_rem * 1e-9, 
         emissions_stay = emissions_stay * 1e-9, 
         total_rem = storage_rem + emissions_rem, 
         total_stay = storage_stay + emissions_stay) %>% 
  ggplot()+
  #geom_hline(aes(yintercept = 0), color = "red", linetype = "dotted", linewidth = 1)+
  geom_line(aes(yr, emissions_stay), linetype = "solid", color = "grey70", linewidth = 1)+
  geom_line(aes(yr, storage_stay), linetype = "solid", color = "grey70", linewidth = 1)+
  geom_line(aes(yr, total_stay), linetype = "dashed", color = "grey70", linewidth = 1)+
  geom_line(aes(yr, storage_rem), linetype = "solid", color = "black", linewidth = 1)+
  geom_line(aes(yr, total_rem), linetype = "dashed", color = "black", linewidth = 1)+
  geom_line(aes(yr, emissions_rem), linetype = "solid", color = "black", linewidth = 1)+
  theme_bw()+
  ggtitle("Veazie Dam")+
  labs(x = "Dam age (yr)", y = expression(Carbon~balance~"(Gg"~CO[2]-eq~")"))+
  theme(axis.text = element_text(size = 14, color = "black"), axis.title = element_text(size =16), plot.title = element_text(size=20, hjust = 0.5))
```

```{r fig 5}
png(paste0("../2-output-data/trajectories_", Sys.Date(), ".png"), width = 10, height = 5, units = "in", res = 300)
plot_grid(veazie_pathways.plt, glines_pathways.plt)
dev.off()
```

# Supplemental information

Emissions before dam removal g C
```{r res emission g C}
all_res %>% 
  filter(flux == "reservoir emissions", flux_unit == "C") %>% 
  dplyr::select(flux, variation, dam, mean_area_mg, ci_lwr_area_mg, ci_upr_area_mg)
```

Burial before dam removal g C
```{r res burial g C}
all_res %>% 
  filter(flux == "reservoir burial", flux_unit == "C") %>% 
  dplyr::select(flux, variation, dam, mean_area_mg, ci_lwr_area_mg, ci_upr_area_mg)
```

Net flux before and after dam removal g C
```{r before after g C}
before_after_df %>% 
  filter(flux_unit == "C") %>% 
  group_by(dam, period) %>% 
  summarize(mean_mg = sum(mean_area_mg), lwr_mg = sum(ci_lwr_area_mg), upr_mg = sum(ci_upr_area_mg))
```

Burp emissions before dam removal g C
```{r}
veazie_scenarios_gC <- all_res %>% 
  filter(dam == "Veazie", period == "burp", flux_unit == "C")%>% 
  mutate(flux_var = paste(flux, variation)) %>% 
  filter(flux_var %in% c("exposed sediment CO2 0.083333333", "exposed sediment CH4 no_pore_0.0833333333333333", "drawdown ebullition 0", 
                         "exposed sediment CO2 5", "exposed sediment CH4 pore_5", "drawdown ebullition 10",  
                         "exposed sediment CO2 30", "exposed sediment CH4 pore_30", "drawdown ebullition 100", "eroded sediment emissions NA"))

veazie_scenarios_rw_gC <- veazie_scenarios_gC %>% slice(10)

veazie_scenarios_gC <- veazie_scenarios_gC %>% bind_rows(veazie_scenarios_rw_gC) %>% bind_rows(veazie_scenarios_rw_gC) %>% 
  mutate(scenario = rep(c("least", "moderate", "most"), 4))

veazie_scenarios_summary_gC <- veazie_scenarios_gC %>% group_by(scenario) %>% summarize(net_mg = sum(mean_area_mg), net_mg_lwr = sum(ci_lwr_area_mg), net_mg_upr = sum(ci_upr_area_mg))

glines_scenarios_gC <- all_res %>% 
  filter(dam == "Glines Canyon", period == "burp", flux_unit == "C")%>% 
  mutate(flux_var = paste(flux, variation)) %>% 
  filter(flux_var %in% c("exposed sediment CO2 0.0833333333333333", "exposed sediment CH4 no_pore_0.0833333333333333", "drawdown ebullition 0", 
                         "exposed sediment CO2 5", "exposed sediment CH4 pore_5", "drawdown ebullition 10",  
                         "exposed sediment CO2 30", "exposed sediment CH4 pore_30", "drawdown ebullition 100", "eroded sediment emissions NA"))

glines_scenarios_rw_gC <- glines_scenarios_gC %>% slice(10)

glines_scenarios_gC <- glines_scenarios_gC %>% bind_rows(glines_scenarios_rw_gC) %>% bind_rows(glines_scenarios_rw_gC) %>% 
  mutate(scenario = rep(c("least", "moderate", "most"), 4))

glines_scenarios_summary_gC <- glines_scenarios_gC %>% group_by(scenario) %>% summarize(net_mg = sum(mean_area_mg), net_mg_lwr = sum(ci_lwr_area_mg), net_mg_upr = sum(ci_upr_area_mg))

veazie_scenarios_summary_gC

veazie_scenarios_gC %>% group_by(scenario) %>% arrange(desc(mean_area_mg)) %>% slice(1)

glines_scenarios_summary_gC

glines_scenarios_gC %>% group_by(scenario) %>% arrange(desc(mean_area_mg)) %>% slice(1)
```

River CO2 emissions after removal gC
```{r}
all_res %>% 
  filter(flux == "river CO2 emissions", flux_unit == "C") %>% 
  dplyr::select(flux, variation, dam, mean_area_mg, ci_lwr_area_mg, ci_upr_area_mg)
```

Terrestrial NEP after removal gC
```{r}
all_res %>% 
  filter(flux == "NEP", flux_unit == "C", variation == 100) %>% 
  dplyr::select(flux, variation, dam, mean_area_mg, ci_lwr_area_mg, ci_upr_area_mg)

all_res %>% filter(dam == "Veazie", flux_unit == "C", period == "build") %>% arrange(desc(mean_area_mg)) %>% dplyr::select(flux, variation, mean_area_mg, ci_lwr_area_mg, ci_upr_area_mg)
```

Figure S1
```{r fig s1}
png(paste0("../2-output-data/fig-s1_", Sys.Date(), ".png"), width = 8, height = 5, units = "in", res = 300)
gen_theme(all_res %>% filter(flux_unit == "CO2-eq", flux == "reservoir emissions") %>% mutate(var = if_else(is.na(variation), "without degassing", variation)) %>% 
  ggplot()+
  geom_point(aes(dam, mean), size = 2)+
  geom_errorbar(aes(x = dam, ymin = ci_lwr, ymax = ci_upr, linetype = var), linewidth = 1, width = 0.5)+
     ylab(expression(g~CO[2-eq]~m^-2~yr^-1))+
    scale_linetype_manual(values = c("dotted", "solid")))
dev.off()
```

Figure S2
```{r Fig S2}
veazie_burp_pub <- all_res %>% filter(dam == "Veazie", period == "burp", flux_unit == "CO2-eq") %>% mutate(flux_var = paste0(flux, variation), var = c("1 month", "1 year", "5 years", "10 years", "30 years", "DNI", "1 month no PE", "1 year no PE", "5 years no PE", "10 years no PE", "30 years no PE", "DNI","1 month + PE", "1 year + PE", "5 years + PE", "10 years + PE", "30 years + PE", "DNI", "x0", "x5", "x10", "x25", "x100", "NA")) %>% filter(var != "DNI") %>% mutate(var_fct = as.factor(var) %>%  fct_relevel(c("1 month", "1 year", "5 years", "10 years", "30 years", "1 month no PE", "1 year no PE", "5 years no PE", "10 years no PE", "30 years no PE", "1 month + PE", "1 year + PE", "5 years + PE", "10 years + PE", "30 years + PE", "x0", "x5", "x10", "x25", "x100", "NA")))

glines_burp_pub <- all_res %>% filter(dam == "Glines Canyon", period == "burp", flux_unit == "CO2-eq") %>% mutate(flux_var = paste0(flux, variation), var = c("1 month", "1 year", "5 years", "10 years", "30 years", "DNI", "1 month no PE", "1 year no PE", "5 years no PE", "10 years no PE", "30 years no PE", "DNI","1 month + PE", "1 year + PE", "5 years + PE", "10 years + PE", "30 years + PE", "DNI", "x0", "x5", "x10", "x25", "x100", "NA")) %>% filter(var != "DNI") %>% mutate(var_fct = as.factor(var) %>%  fct_relevel(c("1 month", "1 year", "5 years", "10 years", "30 years", "1 month no PE", "1 year no PE", "5 years no PE", "10 years no PE", "30 years no PE", "1 month + PE", "1 year + PE", "5 years + PE", "10 years + PE", "30 years + PE", "x0", "x5", "x10", "x25", "x100", "NA")))

veazie_burp_var <- gen_theme(ggplot(veazie_burp_pub, aes(var_fct, mean_area_mg, color = flux))+
  geom_point(size = 2, position = position_dodge(width = 0.7))+
  geom_errorbar(aes(ymin = ci_lwr_area_mg, ymax = ci_upr_area_mg), size = 1, position = position_dodge(width = 0.7), width = 0.5)+
    ggtitle("Veazie Dam")+
  ylab(expression(Mg~CO[2-eq])))+
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title.x = element_blank(), plot.title = element_text(size=20, hjust = 0.5))

glines_burp_var <- gen_theme(ggplot(glines_burp_pub, aes(var_fct, mean_area_mg, color = flux))+
  geom_point(size = 2, position = position_dodge(width = 0.7))+
  geom_errorbar(aes(ymin = ci_lwr_area_mg, ymax = ci_upr_area_mg), size = 1, position = position_dodge(width = 0.7), width = 0.5)+
    ggtitle("Glines Canyon Dam")+
  ylab(expression(Mg~CO[2-eq])))+
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title.x = element_blank(), plot.title = element_text(size=20, hjust = 0.5))

png(paste0("../2-output-data/fig-s2_", Sys.Date(), ".png"), width = 10, height = 10, units = "in", res = 300)
plot_grid(veazie_burp_var, glines_burp_var, ncol = 1)
dev.off()
```

Figure S3
```{r}
veazie_after_var <- gen_theme(all_res %>% filter(dam == "Veazie", flux_unit == "CO2-eq", period == "build") %>% filter(variation %in% c(NA, 100, "offset by canopy angio", "whole trunk angio", "three meters angio", "offset by canopy gymno", "whole trunk gymno", "three meters gymno")) %>% mutate(mean_mg = mean_area *1e-6, lwr_mg = ci_lwr_area * 1e-6, upr_mg = ci_upr_area * 1e-6, flux_var = paste0(flux, variation)) %>% 
  ggplot(aes(flux_var, mean_mg))+
  geom_point(size = 2)+
  geom_hline(yintercept = 0, linetype = "dotted", color = "red", linewidth = 1)+
  ggtitle("Veazie Dam")+
 geom_errorbar(aes(ymin = lwr_mg, ymax = upr_mg), size = 1, position = position_dodge(width = 0.7), width = 0.5)+
  ylab(expression(Mg~CO[2-eq]~yr^-1))+
  scale_x_discrete(labels = c("NEP", expression(river~CH[4]), expression(river~CO[2]), expression(soil~CH[4]), expression(tree~CH[4]~angio~offset),expression(tree~CH[4]~gymno~offset), expression(tree~CH[4]~angio~3~m), expression(tree~CH[4]~gymno~3~m), expression(tree~CH[4]~angio~whole), expression(tree~CH[4]~gymno~whole))))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title.x = element_blank(), plot.title = element_text(size=20, hjust = 0.5))


glines_after_var <- gen_theme(all_res %>% filter(dam == "Glines Canyon", flux_unit == "CO2-eq", period == "build") %>% filter(variation %in% c(NA, 100, "offset by canopy", "whole trunk", "three meters")) %>% mutate(mean_mg = mean_area *1e-6, lwr_mg = ci_lwr_area * 1e-6, upr_mg = ci_upr_area * 1e-6, flux_var = paste0(flux, variation)) %>% ggplot(aes(flux_var, mean_mg))+
  geom_point(size = 2)+
  geom_hline(yintercept = 0, linetype = "dotted", color = "red", linewidth = 1)+
  ggtitle("Glines Canyon Dam")+
 geom_errorbar(aes(ymin = lwr_mg, ymax = upr_mg), size = 1, position = position_dodge(width = 0.7), width = 0.5)+
  ylab(expression(Mg~CO[2-eq]~yr^-1)))+
  scale_x_discrete(labels = c("NEP", expression(river~CH[4]), expression(river~CO[2]), expression(soil~CH[4]), expression(tree~CH[4]~angio~offset), expression(tree~CH[4]~angio~3~m), expression(tree~CH[4]~angio~whole)))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), axis.title.x = element_blank(), plot.title = element_text(size=20, hjust = 0.5))

png(paste0("../2-output-data/fig-s3_", Sys.Date(), ".png"), width = 10, height = 10, units = "in", res = 300)
plot_grid(veazie_after_var, glines_after_var, ncol = 1)
dev.off()
```


# Supplemental analyses not in SI
Imposing constraint on exposed sediment CH4
```{r}
# take the highest reasonable   
max_exp_ch4 <- function(exp_co2){
  return(exp_co2 * (1/44.01)* 1.7 * 16.04 * 34) 
}

# Veazie
veazie_high_med <- veazie %>% filter(flux == "exposed sediment CO2", variation == 10, flux_unit == "CO2-eq") %>% pull(mean) 

max_exp_ch4(veazie_high_med) 

veazie %>% filter(flux == "exposed sediment CH4", variation == "no_pore_30", flux_unit == "CO2-eq")

# Glines
glines_high_med <- glines %>% filter(flux == "exposed sediment CO2", variation == 30, flux_unit == "CO2-eq") %>% pull(mean) 

max_exp_ch4(glines_high_med) 

glines %>% filter(flux == "exposed sediment CH4", variation == "no_pore_30", flux_unit == "CO2-eq")
```

