---
title: "simkins"
author: "L. Naslund"
date: "2024-07-18"
output: html_document
---

The Simkins Dam was the further upstream of two dams removed on the Patapsco River in Maryland.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rnaturalearth)
library(tidyverse)
library(sf)
library(raster)
library(mapview)
library(lme4)
library(MuMIn)
library(readxl)
library(merTools)
library(elevatr)

# required geospatial 
res <- st_read("../1-input-data/1-geospatial/reservoirs/simkins_res.shp")
ws <- st_read("../1-input-data/1-geospatial/watersheds/simkinswatershed.shp") %>% st_transform(st_crs(res)) 
riv <- st_read("../1-input-data/1-geospatial/post-removal/simkins_riv.shp") %>% st_transform(st_crs(res))  

mapview(res)+
  mapview(riv, col.regions = "red")

summary_df <- data.frame(period = character(), flux = character(), variation = character(), flux_unit = character(), annual = character(), area = character(), mean = double(), ci_lwr = double(), ci_upr = double())
```

```{r reservoir and catchment characteristics}
## g-res required inputs
# catchment area (km2) 
sf_use_s2(FALSE)
ws_area <- st_area(ws)
print(paste0("Catchment area (km2): ", round(ws_area * 1e-6, 2)))

# gres required inputs 
ws_pop <- 186078
print(paste0("Population in the catchment: ", ws_pop))

# community wastewater treatment
wwtp <- "secondary"
print(paste0("Wastewater treatment: ", wwtp))

# land cover
simkins_lc <- read.csv("../1-input-data/3-gres/simkins_lc.csv") %>% 
  mutate(gres_class = case_when(
    Legend == "Open Water" ~ "Water Bodies", 
    Legend == "Perennial Snow/Ice" ~ "Permanent Snow/Ice",
    Legend == "Developed, Open Space" ~ "Settlements", 
    Legend == "Developed, Low Intensity" ~ "Settlements", 
    Legend == "Developed, Medium Intensity" ~ "Settlements", 
    Legend == "Developed, High Intensity" ~ "Settlements", 
    Legend == "Barren Land" ~ "Bare Areas",
    Legend == "Deciduous Forest" ~ "Forest", 
    Legend == "Evergreen Forest" ~ "Forest", 
    Legend == "Mixed Forest" ~ "Forest", 
    Legend == "Shrub/Scrub" ~ "Grassland/Shrubland", 
    Legend == "Herbaceuous" ~ "Grassland/Shrubland", 
    Legend == "Hay/Pasture" ~ "Croplands", 
    Legend == "Cultivated Crops" ~ "Croplands",
    Legend == "Woody Wetlands" ~ "Wetlands", 
    Legend == "Emergent Herbaceuous Wetlands" ~ "Wetlands",
  ))

print("Catchment landcover:")
lc <- simkins_lc %>% group_by(gres_class) %>% summarize(perc = sum(percent) %>% round(1)) 
print(lc)

# country
country <- "United States"
print(paste0("Country: ", country))

# coordinates
long <- -76.7683
lat <- 39.2527
print(paste0("Coordinates: ", lat, ",", long))

# reservoir area (km2) 
res_area <- st_area(res) 
print(paste0("Reservoir Area (km2): ", res_area * 1e-6 ))

# mean depth (m) 
mean_depth <- 0.92
print(paste0("Mean Depth (m): ", mean_depth))

# max depth (m) 
max_depth <- 2.7
print(paste0("Max Depth (m): ", max_depth))

# littoral area (%) manually entered b/c g-res defines as <3 m 
lit_area <- 100
print(paste0("Littoral Area (%): ", lit_area))

# thermocline depth (m) Output from g-res 
thermo_depth <- 0.2
print(paste0("Thermocline depth (m): ", thermo_depth))

# soil carbon 
soil_carbon <- 4.36
print(paste0("Soil carbon under impounded area (kgC/m2): ", soil_carbon))

# reservoir wind 
res_wind <- 4.23
print(paste0("Reservoir wind at 50 m: ", res_wind))

# water residence time (yrs) Output from g-res 
wrt <- 0.0002
print(paste0("Water residence time (yrs): ", wrt))

# annual discharge 
q_annual <- 7
print(paste0("Annual Discharge (m3/s): ", q_annual))

# p concentration (ug/L) Output from MD DNR
p_conc <- 50
print(paste0("P Concentration (ug/L): ", p_conc))

# global horizontal radiance 
rad <- 3.85
print(paste0("Global horizontal radiance: ", rad))

# mean temperature per month 
temperature <- data.frame(Jan = 0.1, Feb = 1.5, Mar = 6.1, Apr = 11.7, May = 17.4, June = 22.2, July = 24.8, Aug = 23.9, Sept = 19.9, Oct = 13.6, Nov = 7.9, Dec = 2.4) %>% pivot_longer(cols = Jan:Dec, names_to = "Month", values_to = "Temperature")
print("Temperature per month (C)")
temperature
```


# Before

## Reservoir surface emissions

**G-res tool**

https://131.datatrium.com/fmi/webd/G-res%20Tool?script=ChoiceWebPage&param=Grestool&homeurl=https://g-res.hydropower.org

**Required data**

Must know

- reservoir age
- P concentration (ug/L) or population in the catchment and category of wastewater treatment
- water residence time or mean depth
- reservoir area (km^2)
- reservoir climate zone
- maximum depth
- mean depth or volume

Can be derived from other inputs

- % littoral area
- water residence time

Can be calculated from Google Earth Engine script given a delineated catchment with 
https://code.earthengine.google.com/db5542fcc97469961bfd890123aa55df

- mean annual air temperature (C)
- reservoir surface soil carbon content (kgC m^-2)
- cumulative global horizontal radiance (kWh m^-2 d-1)
- catchment area (km2)
- annual runoff (mm/yr)
- catchment landcover (%)
- k factor (soil erodibility factor)
- mean basin slope

Can delineate catchments in US reservoirs using the StreamStats tool 
https://streamstats.usgs.gov/ss/ 

```{r surface emissions}
# output from g-res: emissions in g CO2 eq/m2/yr by pathway
res_co2_diff_eq <- 74
res_ch4_diff_eq <- 402 * .49
res_ch4_eb_eq <- 402 * 0.51
res_ch4_degas_eq <- 0

# output from g-res: net GHG footprint and 95% CI in g CO2 eq/m2/yr
res_net_eq <- 806
res_net_2.5_eq <- 753
res_net_97.5_eq <- 867

res_net_interval <- max(c(res_net_97.5_eq-res_net_eq, res_net_eq - res_net_2.5_eq))

# calculate emissions and CI in g CO2 eq/m2/yr 
res_eq <- res_co2_diff_eq+res_ch4_diff_eq+res_ch4_eb_eq

res_eq_2.5 <- res_eq - res_net_interval
res_eq_97.5 <-  res_eq + res_net_interval

# calculate emissions and CI in g C/m2/yr
res_g <- (res_co2_diff_eq * (12.01/44.01)) + ((res_ch4_diff_eq+res_ch4_eb_eq + res_ch4_degas_eq) *(12.01/16.04) * (1/34))

# to calculate CI in g C, I am assuming gases contribute to the CI in proportion to the CO2-eq contribution 
res_g_interval <- res_net_interval*(res_co2_diff_eq/res_eq)*(12.01/44.01) + res_net_interval*((res_ch4_diff_eq + res_ch4_eb_eq + res_ch4_degas_eq)/res_eq)*(1/34) * (12.01/16.04)

res_g_2.5 <- res_g - res_g_interval
res_g_97.5 <- res_g + res_g_interval

# summary data frame
res_emissions_summary <- data.frame(period = c("before"), flux = c("reservoir emissions"), variation = NA, flux_unit = c("CO2-eq", "C"), annual = "yes", area = "reservoir", mean = c(res_eq, res_g), ci_lwr = c(res_eq_2.5, res_g_2.5), ci_upr = c(res_eq_97.5, res_g_97.5))

res_emissions_summary
```
res

## Reservoir carbon burial
```{r carbon burial, warnings = FALSE}
# sedimentation model
sed_df <- read.csv("../1-input-data/6-clow-data/clow_predictors.csv") %>% 
    mutate(sed_rate_mm_yr = `Sed.Rate..mm.yr.`,
        surface_area_m2 = as.numeric(Surface_Area_acres) * 4046.86,
        sed_rate_m3_yr = sed_rate_mm_yr * 1e-3 * surface_area_m2, 
         log_sed_rate_m3_yr = log10(sed_rate_m3_yr), 
         log_surface_area_m2 = log10(surface_area_m2), 
         crop_perc = crops_calc/100, 
         forest_perc = forest_calc/100)

sed.mod <- lm(log_sed_rate_m3_yr ~ log_surface_area_m2 + slope_calc + forest_perc + crop_perc, data = sed_df)

summary(sed.mod)

simkins_sed_pred <- read.csv("../1-input-data/6-clow-data/simkins_clow.csv") %>% 
  mutate(log_surface_area_m2 = log10(res_area %>% as.numeric()), 
         crop_perc = crops/100, 
         forest_perc = forest/100) %>% 
  rename("slope_calc" = "slope")

predict(sed.mod, simkins_sed_pred, interval = "prediction")

simkins_sed_rate <- 10^(predict(sed.mod, simkins_sed_pred, interval = "prediction")[1])
simkins_sed_rate_sd <- (10^(predict(sed.mod, simkins_sed_pred, interval = "prediction")[1]) - 10^(predict(sed.mod, simkins_sed_pred, interval = "prediction")[2]))/1.96


# sediment c content model
c_content <- read.csv("../1-input-data/6-clow-data/c_content.csv") %>% 
  mutate(LAKE_ORIGIN_renamed = as.factor(LAKE_ORIGIN_renamed)) %>% 
  dplyr::select(sediment_C_pct, SOC_0_5, Wetlands90, log_Wetlands90, Barren30, KFactor, 
                log_WaterBody_Area_m2, log_Lake_Area_m2_EPA, LAKE_ORIGIN_renamed) %>% mutate(wetland_si = log10(Wetlands90 + 1))

c.mod <- lm(sediment_C_pct ~ SOC_0_5 + wetland_si + Barren30 + KFactor + log_Lake_Area_m2_EPA + LAKE_ORIGIN_renamed, data = c_content)

simkins_c_pred <- simkins_sed_pred %>% 
  rename("SOC_0_5" = "om", "KFactor" = "k_factor") %>% 
  mutate(wetland_si = log10((wetland/100)+1), 
         Barren30 = barren/100, 
         log_Lake_Area_m2_EPA = log10(res_area), 
         LAKE_ORIGIN_renamed = "RESERVOIR")

predict(c.mod, simkins_c_pred, interval = "prediction")

simkins_c_cont <- predict(c.mod, simkins_c_pred, interval = "prediction")[1] * 0.01 # covert % to proportion
simkins_c_cont_sd <- ((predict(c.mod, simkins_c_pred, interval = "prediction")[1] - predict(c.mod, simkins_c_pred, interval = "prediction")[2])/1.96)*0.01

b = (simkins_c_cont_sd)^2/simkins_c_cont
a = (simkins_c_cont)^2/(simkins_c_cont_sd)^2
  
# dry bulk density
# simkins has a dry bulk density measured so don't need to use Dean and Gorham
# Collins et al. 2017 cited mean 1.43 g/cm3 and range of 1.29-1.53. 
simkins_dbd <- 1.43 * 1e6 #g/m3
simkins_dbd_sd <- (max((1.53-1.43), (1.43-1.29))/1.96) * 1e6 # g/m3

# burial efficiency
be <- read.csv("../1-input-data/6-clow-data/sed-rate_burial-efficiency.csv", header = FALSE) %>% rename("log_sed_rate_mm_yr" = "V1", "oc_be" = "V2")

be.mod <- lm(oc_be ~ log_sed_rate_mm_yr, data = be)

simkins_burial <- c()

for(i in 1:10000){
  print(i)
  
  # sample sedimentation rate and OC content
  temp_sed_rate <- rnorm(1, simkins_sed_rate, simkins_sed_rate_sd)
  temp_oc <- rgamma(1, shape = a, scale = b)
  
  # get dbd
  temp_dbd <- rnorm(1, simkins_dbd, simkins_dbd_sd)
  
  # get burial efficiency
  temp_be_df <- data.frame(log_sed_rate_mm_yr = log10(temp_sed_rate * (1/as.numeric(res_area)) * 1e3))
  
  temp_be_distr <- predict(be.mod, temp_be_df, interval = "prediction")[1] * 0.01
  temp_be_distr_sd <- ((predict(be.mod, temp_be_df, interval = "prediction")[1] * 0.01)-predict(be.mod, temp_be_df, interval = "prediction")[2] * 0.01)/1.96
  
  d = (temp_be_distr_sd)^2/temp_be_distr
  c = (temp_be_distr)^2/(temp_be_distr_sd)^2
  
  temp_be <- rgamma(1, shape = c, scale = d)
  
  temp <- temp_sed_rate * temp_dbd * temp_oc * temp_be * (1/as.numeric(res_area))  
  
  simkins_burial <- c(simkins_burial, temp)

}

simkins_burial <- sort(simkins_burial) # sort removes NAs, by assigning prediction interval to a normal distribution, generating some few negative values of sedimentation rate, sort here functionally truncates the distribution at 0.  
vec_length <- length(simkins_burial)

res_burial_summary <- data.frame(period = c("before"), flux = c("reservoir burial"), variation = NA, flux_unit = c("CO2-eq", "C"), annual = "yes", area = "reservoir", mean = c(-(simkins_burial[round(vec_length * 0.5)] * (44.01/12.01)), -simkins_burial[round(vec_length * 0.5)]), ci_lwr = c(-(simkins_burial[round(vec_length * 0.975)] * (44.01/12.01)), -simkins_burial[round(vec_length * 0.975)]), ci_upr = c(-(simkins_burial[round(vec_length * 0.025)] * (44.01/12.01)), -simkins_burial[round(vec_length * 0.025)]))

res_burial_summary

# calculations for fig 5
simkins_sed_fig5 <- data.frame(oc = simkins_c_cont, dbd = simkins_dbd, area = res_area)

write.csv(simkins_sed_fig5, "../2-output-data/simkins_sed_fig5.csv", row.names = FALSE)
```

# Burp

## Exposed sediment CO2 emissions 
```{r exposed sed co2}
res_exp_sed_co2 <- read.csv("../2-output-data/exp_sed_co2_sampled.csv")

res_sed_co2_summary <- data.frame(period = c("burp"), flux = c("exposed sediment CO2"), variation = res_exp_sed_co2$y %>% as.character(), flux_unit = c(rep("CO2-eq", 6), rep("C", 6)), annual = "no", area = "exposed sediment area", mean = c(res_exp_sed_co2$mean_co2, res_exp_sed_co2$mean_c), ci_lwr = c(res_exp_sed_co2$lwr_co2, res_exp_sed_co2$lwr_c), ci_upr = c(res_exp_sed_co2$upr_co2, res_exp_sed_co2$upr_c))
```

## Exposed sediment CH4 emissions 
```{r exposed sed ch4}
thresh_0.01 <- read.csv("../2-output-data/glines_exp_sed_ch4.csv")

pore_co2_eq <- 8102.2 * 1e-3 * (16.04/12.01)* 34
pore_c <- 8102.2 * 1e-3

res_sed_ch4_summary <- data.frame(period = c("burp"), flux = c("exposed sediment CH4"), variation = c(rep(paste0("no_pore_", thresh_0.01$y), 2), rep(paste0("pore_", thresh_0.01$y), 2)), flux_unit = rep(c(rep("CO2-eq", 6), rep("C", 6)), 2), annual = "no", area = "exposed sediment area", mean = c(thresh_0.01$mean_co2,  thresh_0.01$mean_c, thresh_0.01$mean_co2 + pore_co2_eq, thresh_0.01$mean_c + pore_c), ci_lwr = c(thresh_0.01$lwr_co2,  thresh_0.01$lwr_c, thresh_0.01$lwr_co2 + pore_co2_eq, thresh_0.01$lwr_c + pore_c), ci_upr = c(thresh_0.01$upr_co2, thresh_0.01$upr_c, thresh_0.01$upr_co2 + pore_co2_eq, thresh_0.01$upr_c + pore_c))
```

## Drawdown ebullition 
```{r drawdown eb}
res_ch4_eb_eq_2.5 <- res_ch4_eb_eq - (res_net_interval * (res_ch4_eb_eq/res_eq))
res_ch4_eb_eq_97.5 <- res_ch4_eb_eq + (res_net_interval * (res_ch4_eb_eq/res_eq))

drawdown_eb_prop <- c(0, 5, 10, 25, 100)

res_ch4_eb_g <- res_ch4_eb_eq * (1/34) * (12.01/16.04)
res_ch4_eb_g_2.5 <- res_ch4_eb_eq_2.5 * (1/34) * (12.01/16.04)
res_ch4_eb_g_97.5 <- res_ch4_eb_eq_97.5 * (1/34) * (12.01/16.04)

res_drawdown_eb_summary <- data.frame(period = c("burp"), flux = c("drawdown ebullition"), variation = drawdown_eb_prop %>% as.character(), flux_unit = c(rep("CO2-eq",5), rep("C", 5)), annual = "no", area = "reservoir", mean = c(res_ch4_eb_eq * drawdown_eb_prop, res_ch4_eb_g * drawdown_eb_prop) , ci_lwr = c(res_ch4_eb_eq_2.5 * drawdown_eb_prop, res_ch4_eb_g_2.5 * drawdown_eb_prop) , ci_upr = c(res_ch4_eb_eq_97.5 * drawdown_eb_prop, res_ch4_eb_g_97.5 * drawdown_eb_prop))
```

## Eroded sediment emissions 
```{r eroded sed}
# Collins et al. write that a combination of the CR Environmental Report and post-removal survey suggest that the combined volume of mobilized sediment from Veazie and Great Works dam was 15,000 m3 or smaller
# Fit a uniform distribution between 0 and 15,000 m3 

b = (simkins_c_cont_sd)^2/simkins_c_cont
a = (simkins_c_cont)^2/(simkins_c_cont_sd)^2

# assumming between 0 and 100% oxidized
min_ox_trans = 0 
max_ox_trans = 1

eroded_vec <- c()
for(i in 1:10000){
  print(i)
  
  # sample OC content
  temp_oc <- rgamma(1, shape = a, scale = b)
  
  temp_dbd <- rnorm(1, simkins_dbd, simkins_dbd_sd)
  
  # parameters for g of sediment
  res_erode_g_mean <- 56350 * temp_dbd # g
  res_erode_g_sd <- (31000 * 1e6)/1.96
  
  temp <- (rnorm(1, mean = res_erode_g_mean, sd = res_erode_g_sd) * temp_oc * runif(1, min = min_ox_trans, max = max_ox_trans))/(res_area)
  eroded_vec <- c(eroded_vec, temp)
}

eroded_vec <- sort(eroded_vec)

res_erode_summary <- data.frame(period = c("burp"), flux = c("eroded sediment emissions"), variation = NA, flux_unit = c("CO2-eq", "C"), annual = "no", area = "reservoir", mean = c(eroded_vec[5000] * (44.01/12.01), eroded_vec[5000]), ci_lwr = c(eroded_vec[250] * (44.01/12.01), eroded_vec[250]), ci_upr = c(eroded_vec[9750] * (44.01/12.01), eroded_vec[9750]))
```

# Build

## River surface CO2 emissions 
```{r river co2 emissions}
# Discharge is 7 m3/s which puts it in the 1.0-10 m3/s bin
# Values are reported as g C m-2 d-

# Function to calculate the sum of squared differences between observed and target quantiles

gamma_params_diff <- function(params, quantiles_target) {
  shape <- params[1]
  scale <- params[2]
  
  quantiles_observed <- qgamma(c(0.25, 0.5, 0.75), shape = shape, scale = scale)
  
  sum((quantiles_observed - quantiles_target)^2)
}

# Target quantiles
quantiles_target <- c(2.18, 3.57, 5.74)

# Initial guess for shape and scale parameters
initial_params <- c(shape = 1, scale = 1)

# Optimize the parameters
optimized_params <- optim(par = initial_params, fn = gamma_params_diff, quantiles_target = quantiles_target)

# Extract optimized parameters
shape_optimized <- optimized_params$par[1]
scale_optimized <- optimized_params$par[2]

# Check the quantiles with the optimized parameters
quantiles_optimized <- qgamma(c(0.25, 0.5, 0.75), shape = shape_optimized, scale = scale_optimized)
cat("Optimized Quantiles:", quantiles_optimized, "\n")

x <- seq(0, 10, by = 0.1)
y <- dgamma(x, shape = shape_optimized, scale = scale_optimized)

plot(x, y)

riv_co2_vec <- sort(rgamma(10000, shape = shape_optimized, scale = scale_optimized)) *365

riv_co2_summary <- data.frame(period = c("build"), flux = c("river CO2 emissions"), variation = NA, flux_unit = c("CO2-eq", "C"), annual = "yes", area = "river", mean = c(riv_co2_vec[5000] * (44.01/12.01), riv_co2_vec[5000]), ci_lwr = c(riv_co2_vec[250] * (44.01/12.01), riv_co2_vec[250]), ci_upr = c(riv_co2_vec[9750] * (44.01/12.01), riv_co2_vec[9750]))
```

## River surface CH4 emissions 
```{r river ch4 emissions}
ch4_conc <- read.csv("../1-input-data/5-ch4-fluxes/GRiMe_concentrations_v2.csv") 
ch4_flux <- read.csv("../1-input-data/5-ch4-fluxes/GRiMe_fluxes_v2.csv")
ch4_flux_sites <- read.csv("../1-input-data/5-ch4-fluxes/GRiMe_sites_v2.csv")
ch4_source <- read.csv("../1-input-data/5-ch4-fluxes/GRiMe_sources_v2.csv")

usa <- ne_countries(country = "united states of america")

ch4_flux_sf <- ch4_flux %>% 
  left_join(ch4_flux_sites, by = c("Source_ID", "Site_ID", "Site_Name")) %>% 
  left_join(ch4_conc %>% 
              filter(FluxYesNo == "Yes") %>% 
              rename("Flux_Name" = "Conc_Name"), 
            by = c("Source_ID", "Site_ID", "Site_Name", "Flux_Name")) %>% 
  filter(is.na(Longitude)==FALSE, is.na(Channel_type)==TRUE) %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326) 

# filter sites in us
ch4_flux_sites_usa <- st_intersection(usa, ch4_flux_sf) %>% dplyr::select(Source_ID:geometry)
mapview(ch4_flux_sites_usa)

# Does it make sense to bin by discharge just because Hotckiss et al. 2015 did if discharge is not explanatory?
ch4_flux_sites_usa %>% ggplot(aes(Q, Diffusive_CH4_Flux_Mean))+geom_point()
ch4_flux_sites_usa %>% filter(Q < 500 & Diffusive_CH4_Flux_Mean<45) %>% ggplot(aes(Q, Diffusive_CH4_Flux_Mean))+geom_point()

ch4_flux_sites_usa <- ch4_flux_sites_usa %>% mutate(q_bin = case_when(
  (Q>=0 & Q<0.01) ~ "0-0.01",
  (Q>0.01 & Q<0.1) ~ "0.01-0.1",
  (Q>0.1 & Q<1) ~ "0.1-1",
  (Q>1 & Q<10) ~ "1-10",
  (Q>10 & Q<100) ~ "10-100",
  (Q > 100) ~ ">100"
)) 

# summary of diffusive CH4 by bin
ch4_flux_sites_usa %>% filter(is.na(Diffusive_CH4_Flux_Mean)==FALSE) %>% group_by(q_bin) %>% summarize(count = n())

# no measurements of ebullition in the Q range so assuming it is zero. This is another caveat we will need to address
ch4_flux_sites_usa %>% filter(is.na(Eb_CH4_Flux_Mean)==FALSE) %>% group_by(q_bin) %>% summarize(count = n())


ch4_flux_vec <- sort(sample(ch4_flux_sites_usa %>% filter(q_bin== "1-10") %>% pull(Diffusive_CH4_Flux_Mean), 10000, replace = TRUE))
ch4_flux_vec_gCH4 <- ch4_flux_vec * 365 * 1e-3 * 16.04
ch4_flux_vec_gCO2eq <- ch4_flux_vec_gCH4 * 34

# CO2 flux
co2_vec <- sort(sample(ch4_flux_sites_usa %>% filter(is.na(CO2_Flux_Mean)==FALSE)  %>% filter(q_bin== "1-10") %>% pull(CO2_Flux_Mean), 10000, replace = TRUE))* 365 * 1e-3 * 44.01
co2_vec[250]
co2_vec[5000]
co2_vec[9750]

riv_ch4_summary <- data.frame(period = c("build"), flux = c("river CH4 emissions"), variation = NA, flux_unit = c("CO2-eq", "C"), annual = "yes", area = "river", mean = c(ch4_flux_vec_gCO2eq[5000], ch4_flux_vec_gCH4[5000] * (12.01 / 16.04)), ci_lwr = c(ch4_flux_vec_gCO2eq[250], ch4_flux_vec_gCH4[250] * (12.01 / 16.04)), ci_upr = c(ch4_flux_vec_gCO2eq[9750], ch4_flux_vec_gCH4[9750] * (12.01 / 16.04)))
```

## NEP 
```{r nee}
f_age <- function(a, b, e, age){
  return(e + a * (1-exp(b*age)))
}

df <- read.csv("../1-input-data/8-besnard-data/besnard_2018.csv") %>% mutate(age_func = f_age(263.2, -0.19, 0, Age)) %>% mutate(nutrient = as.factor(`NA.`))

# estimating NEP assuming that it is zero before the model predicts positive values
nep_mod <- lm(NEP ~ MAT + Age + age_func + nutrient, data = df)
summary(nep_mod)

nep_new <- data.frame(MAT = mean(temperature$Temperature), Age = seq(0, 100, by = 1), age_func = f_age(263.2, -0.19, 0, seq(0, 100, by = 1)), nutrient =as.factor("L"))
nep_pred <- predict(nep_mod, nep_new, interval = "prediction") %>% as.data.frame()

# becomes 0 a little before 5.7 years in so going to call it 6

nep_df <- data.frame(years = double(), mean_gC = double(), lwr_gC = double(), upr_gC = double())

year_vec <- c(25,50,100)
for(i in 1:length(year_vec)){
  yr <- year_vec[i]
  mean_gC <- integrate(function(x) as.data.frame(predict(nep_mod, newdata = data.frame(MAT = mean(temperature$Temperature), Age = x, age_func = f_age(263.2, -0.19, 0, x), nutrient =as.factor("L")), interval = "prediction", level = 0.95))$fit, lower = 6, upper = yr)$value
  lwr_gC <- -(integrate(function(x) -(as.data.frame(predict(nep_mod, newdata = data.frame(MAT = mean(temperature$Temperature), Age = x, age_func = f_age(263.2, -0.19, 0, x), nutrient =as.factor("L")), interval = "prediction", level = 0.95))$lwr), lower = 6, upper = yr))$value
  upr_gC <- integrate(function(x) as.data.frame(predict(nep_mod, newdata = data.frame(MAT = mean(temperature$Temperature), Age = x, age_func = f_age(263.2, -0.19, 0, x), nutrient =as.factor("L")), interval = "prediction", level = 0.95))$upr, lower = 6, upper = yr)$value
  
  temporary <- data.frame(years = yr, mean_gC = -mean_gC, lwr_gC = -upr_gC, upr_gC = -lwr_gC)
  
  nep_df <- nep_df %>% bind_rows(temporary)
}

# integrated carbon
nep_df <- nep_df %>% mutate(mean_gCO2 = mean_gC * (44.01/12.01), lwr_gCO2 = lwr_gC * (44.01/12.01), upr_gCO2 = upr_gC * (44.01/12.01))

# average annual rates
nep_df_rates <- nep_df %>% mutate(across(mean_gC:upr_gCO2, ~ .x / years))

nep_df_rates

res_nep_summary <- data.frame(period = c("build"), flux = c("NEP"), variation = nep_df_rates$years %>% as.character(), flux_unit = c(rep("CO2-eq", 3), rep("C", 3)), annual = "yes", area = "exposed sediment area", mean = c(nep_df_rates$mean_gCO2, nep_df_rates$mean_gC), ci_lwr = c(nep_df_rates$lwr_gCO2, nep_df_rates$lwr_gC), ci_upr = c(nep_df_rates$upr_gCO2, nep_df_rates$upr_gC))

res_nep_summary
```

## Soil CH4 emissions 
```{r soil ch4}
soil <- read.csv("../1-input-data/7-gattica-data/MethaneSoilForest.csv") %>% mutate(MAP_sq = MAP^2, MAT_sq = MAT^2) %>% filter(complete.cases(.))

soil_scale <- soil %>% mutate(across(c(MAP:soilpH, MAP_sq, MAT_sq), scale))

soil_mod <- lmer(Methane ~ MAP*MAT + soilpH + SOC + BulkDens + MAT_sq + SoilNit + MAP_sq + SoilSand + (1 + MAP*MAT + soilpH + SOC + BulkDens + MAT_sq + SoilNit + MAP_sq + SoilSand| Biome), data = soil_scale)

# I am not getting the same AICc
summary(soil_mod)
AICc(soil_mod, REML = FALSE)
AICc(soil_mod, REML = TRUE)

# need to get the predictors for sites
# MAP, MAT, soilpH, BulkDens, SoilNit, Soil Sand

simkins_soil <- read.csv("../1-input-data/7-gattica-data/simkins_soil.csv")

# going to have to manually scale                                                         
soil_data_new <- data.frame(MAP = (simkins_soil$simkins_map- mean(soil$MAP))/sd(soil$MAP),
                            MAT = (simkins_soil$simkins_mat - mean(soil$MAT))/sd(soil$MAT), 
                            soilpH = (simkins_soil$simkins_ph - mean(soil$soilpH))/sd(soil$soilpH), 
                            SOC = (simkins_soil$simkins_oc - mean(soil$SOC))/sd(soil$SOC), 
                            BulkDens = (simkins_soil$simkins_bd - mean(soil$BulkDens))/sd(soil$BulkDens), 
                            SoilNit = (simkins_soil$simkins_tn - mean(soil$SoilNit))/sd(soil$SoilNit), 
                            SoilSand = (simkins_soil$simkins_sand - mean(soil$SoilSand))/sd(soil$SoilSand), 
                            MAT_sq = (simkins_soil$simkins_mat^2 - mean(soil$MAT_sq))/sd(soil$MAT_sq), 
                            MAP_sq = (simkins_soil$simkins_map^2 - mean(soil$MAP_sq))/sd(soil$MAP_sq), 
                            Biome = "Temperate")


# returns values in kg Ch4 /ha/yr
soil_ch4_out <- predictInterval(soil_mod, soil_data_new, n.sims = 10000, level = 0.95) %>% as.data.frame()

soil_ch4_out <- soil_ch4_out %>% mutate(soil_c = fit * 0.1 * (12.01/16.04), 
                        soil_c_2.5 = lwr * 0.1 * (12.01/16.04), 
                        soil_c_97.5 = upr * 0.1 * (12.01/16.04), 
                        soil_co2_eq = fit * 0.1 * 34, 
                        soil_co2_eq_2.5 = lwr * 0.1 * 34, 
                        soil_co2_eq_97.5 = upr * 0.1 * 34)  
  
soil_ch4_out

res_soil_ch4_summary <- data.frame(period = c("build"), flux = c("soil CH4 emissions"), variation = NA, flux_unit = c("CO2-eq", "C"), annual = "yes", area = "exposed sediment area", mean = c(soil_ch4_out$soil_co2_eq, soil_ch4_out$soil_c), ci_lwr = c(soil_ch4_out$soil_co2_eq_2.5, soil_ch4_out$soil_c_2.5), ci_upr = c(soil_ch4_out$soil_co2_eq_97.5, soil_ch4_out$soil_c_97.5))

res_soil_ch4_summary
```

## Tree CH4 emissions 
```{r fvs inputs}
simkins_dem <- elevatr::get_elev_raster(res, z=14, expand = 2000)
simkins_masked <- terra::mask(simkins_dem, res)

elev <- mean(getValues(simkins_masked), na.rm = T) * 3.28084
elev

simkins_slope <- terrain(simkins_dem, opt = "slope")
simkins_slope_masked <- terra::mask(simkins_slope, res)

plot(simkins_slope_masked)

range(getValues(simkins_slope_masked), na.rm = T)
range(getValues(simkins_slope), na.rm = T)

slope <- mean(getValues(simkins_slope_masked), na.rm = T)
slope

simkins_aspect <- terrain(simkins_dem, opt = "aspect")
simkins_aspect_masked <- terra::mask(simkins_aspect, res)

aspect <- mean(getValues(simkins_aspect_masked), na.rm = T)
aspect

res %>% st_transform(crs = 4326) %>% st_centroid() %>% st_coordinates()
```

```{r tree ch4 angio, warnings = FALSE}
trees_k09_be <- read_excel("../1-input-data/2-fvs/simkins-output-boxelder-k09.xlsx")
trees_k11_be <- read_excel("../1-input-data/2-fvs/simkins-output-boxelder-k11.xlsx")

# DBH in inches Ht in ft convert to m2
sa_k09_be <- trees_k09_be %>% 
  mutate(surface_area = (DBH/12)*pi *EstHt *0.092903 * TPA, surface_area_3m = (DBH/12)*pi * 9.84 *0.092903 * TPA) %>% 
  group_by(StandID, Year) %>% summarize(total_surface_area = sum(surface_area), total_surface_area_3m = sum(surface_area_3m)) %>% 
  mutate(habitat = "wetland", tree_type = "angiosperm") %>% ungroup()

sa_k11_be <- trees_k11_be %>% 
  mutate(surface_area = (DBH/12)*pi *EstHt *0.092903 * TPA, surface_area_3m = (DBH/12)*pi * 9.84 *0.092903 * TPA) %>% 
  group_by(StandID, Year) %>% summarize(total_surface_area = sum(surface_area), total_surface_area_3m = sum(surface_area_3m)) %>% 
  mutate(habitat = "wetland", tree_type = "angiosperm") %>% ungroup()


# Pitz et al. 2018
# ug CH4/m2/hr -> g/m2/yr
angio_wet_tree_ch4 <- 567.9 * 8760 *1e-6
angio_wet_tree_ch4_sd <- 523.5 * 8760 *1e-6

# m2 in an acre
plot_area <- 4046.86

angio_ch4_df <- NULL

for(i in 1:10000){
  print(i)

  ch4_rate <- rnorm(1, angio_wet_tree_ch4, angio_wet_tree_ch4_sd)
  tree_sa <- sample(c(1, 2), size = 1, replace = TRUE)

  if(tree_sa == 1){
    temporary <- sa_k09_be %>%
      mutate(ch4_all = total_surface_area * ch4_rate,
             ch4_3m = total_surface_area_3m * ch4_rate) %>%
      mutate(ch4_all_rate = ch4_all/plot_area, ch4_3m_rate = ch4_3m/plot_area) %>%
      summarize(mean_ch4_all = mean(ch4_all_rate),
        mean_ch4_3m = mean(ch4_3m_rate),
        ch4_all_rate_co2 = mean_ch4_all * 34,
        ch4_3m_rate_co2 = mean_ch4_3m * 34)
  }

  if(tree_sa == 2){
      temporary <- sa_k11_be %>%
      mutate(ch4_all = total_surface_area * ch4_rate,
             ch4_3m = total_surface_area_3m * ch4_rate) %>%
      mutate(ch4_all_rate = ch4_all/plot_area, ch4_3m_rate = ch4_3m/plot_area) %>%
      summarize(mean_ch4_all = mean(ch4_all_rate),
        mean_ch4_3m = mean(ch4_3m_rate),
        ch4_all_rate_co2 = mean_ch4_all * 34,
        ch4_3m_rate_co2 = mean_ch4_3m * 34)
  }

  angio_ch4_df <- angio_ch4_df %>% bind_rows(temporary)
}


write.csv(angio_ch4_df, "../2-output-data/simkins_angio_ch4_emissions_3m.csv")
```

```{r angio summary}
angio_ch4_df <- read.csv("../2-output-data/simkins_angio_ch4_emissions_3m.csv")

angio_ch4_all_mean <- (angio_ch4_df$mean_ch4_all %>% sort())[5000] * (12.01/16.04)
angio_ch4_all_2.5 <- (angio_ch4_df$mean_ch4_all %>% sort())[250] * (12.01/16.04)
angio_ch4_all_97.5 <- (angio_ch4_df$mean_ch4_all %>% sort())[9750] * (12.01/16.04)

angio_ch4_3m_mean <- (angio_ch4_df$mean_ch4_3m %>% sort())[5000] * (12.01/16.04)
angio_ch4_3m_2.5 <- (angio_ch4_df$mean_ch4_3m %>% sort())[250] * (12.01/16.04)
angio_ch4_3m_97.5 <- (angio_ch4_df$mean_ch4_3m %>% sort())[9750] * (12.01/16.04)

angio_ch4_all_mean_co2 <- (angio_ch4_df$ch4_all_rate_co2 %>% sort())[5000] * (12.01/16.04)
angio_ch4_all_2.5_co2 <- (angio_ch4_df$ch4_all_rate_co2 %>% sort())[250] * (12.01/16.04)
angio_ch4_all_97.5_co2 <- (angio_ch4_df$ch4_all_rate_co2 %>% sort())[9750] * (12.01/16.04)

angio_ch4_3m_mean_co2 <- (angio_ch4_df$ch4_3m_rate_co2 %>% sort())[5000] * (12.01/16.04)
angio_ch4_3m_2.5_co2 <- (angio_ch4_df$ch4_3m_rate_co2 %>% sort())[250] * (12.01/16.04)
angio_ch4_3m_97.5_co2 <- (angio_ch4_df$ch4_3m_rate_co2 %>% sort())[9750] * (12.01/16.04)
```

```{r tree ch4 summary}
res_tree_ch4_summary <- data.frame(period = c("build"),
                                   flux = c("tree CH4 emissions"),
                                   variation = c("offset by canopy angio", 
                                                 "offset by canopy angio", 
                                                 "whole trunk angio", 
                                                 "whole trunk angio", 
                                                 "three meters angio", 
                                                 "three meters angio"),
                                   flux_unit = rep(c("CO2-eq", "C"), 3),
                                   annual = "yes",
                                   area = "exposed sediment area",
                                   mean = c(0, 
                                            0, 
                                            angio_ch4_all_mean_co2, 
                                            angio_ch4_all_mean, 
                                            angio_ch4_3m_mean_co2, 
                                            angio_ch4_3m_mean),
                                   ci_lwr = c(0, 
                                            0, 
                                            angio_ch4_all_2.5_co2, 
                                            angio_ch4_all_2.5, 
                                            angio_ch4_3m_2.5_co2, 
                                            angio_ch4_3m_2.5),
                                   ci_upr = c(0, 
                                            0, 
                                            angio_ch4_all_97.5_co2, 
                                            angio_ch4_all_97.5, 
                                            angio_ch4_3m_97.5_co2, 
                                            angio_ch4_3m_97.5))

res_tree_ch4_summary

```

# Summary
```{r create summary dfs}
res_area <- st_area(res) %>% as.numeric()
riv_area <- st_area(riv) %>% as.numeric()

area_df <- data.frame(area = c("reservoir", "exposed sediment area", "river"), 
                      area_m2 = c(res_area, (res_area - riv_area), riv_area))

# negative values are for uptake or burial positive values are for emissions

summary_df <- bind_rows(res_emissions_summary, res_burial_summary, res_sed_co2_summary, res_sed_ch4_summary, res_drawdown_eb_summary, res_erode_summary, riv_co2_summary, riv_ch4_summary, res_nep_summary, res_soil_ch4_summary, res_tree_ch4_summary) %>% left_join(area_df, by = "area") %>% mutate(mean_area = mean * area_m2, ci_lwr_area = ci_lwr * area_m2, ci_upr_area = ci_upr * area_m2) %>% mutate(flux_fct = fct_relevel(flux, c("reservoir emissions", "reservoir burial", "exposed sediment CO2", "exposed sediment CH4", "drawdown ebullition", "eroded sediment emissions", "river CO2 emissions", "river CH4 emissions", "NEP", "soil CH4 emissions", "tree CH4 emissions")))


write.csv(summary_df, "../2-output-data/simkins_summary_2024-07-25.csv", row.names = FALSE)
```
