---
title: "veazie"
author: "L. Naslund"
date: "2024-01-08"
output: html_document
---

The Veazie Dam was the further downstream of two dams removed on the Penobscot River in coastal central Maine.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rnaturalearth)
library(tidyverse)
library(sf)
library(raster)
library(mapview)
library(lme4)
library(MuMIn)

# required geospatial 
res <- st_read("../1-input-data/1-geospatial/reservoirs/veazie_res.shp")
ws <- st_read("../1-input-data/1-geospatial/watersheds/veaziewatershed.shp") %>% st_transform(st_crs(res))

st_bbox(ws %>% st_transform(4326))

summary_df <- data.frame(period = character(), flux = character(), variation = character(), flux_unit = character(), annual = character(), area = character(), mean = double(), ci_lwr = double(), ci_upr = double())

# https://soilgrids.org/
# 2108.599973611574 dg/kg

# I think you are going to need bulk density too

```

```{r reservoir and catchment characteristics}
## g-res required inputs
# catchment area (km2)
sf_use_s2(FALSE)
ws_area <- st_area(ws)
print(paste0("Catchment area (km2): ", round(ws_area * 1e-6, 2)))

# gres required inputs
ws_pop <- 79672
print(paste0("Population in the catchment: ", ws_pop))

# community wastewater treatment
wwtp <- "secondary"
print(paste0("Wastewater treatment: ", wwtp))

# land cover
veazie_lc <- read.csv("../1-input-data/3-gres/veazie_lc.csv") %>% 
  mutate(gres_class = case_when(
    Legend == "Open Water" ~ "Water Bodies", 
    Legend == "Perennial Snow/Ice" ~ "Permanent Snow/Ice",
    Legend == "Developed, Open Space" ~ "Settlements", 
    Legend == "Developed, Low Intensity" ~ "Settlements", 
    Legend == "Developed, Medium Intensity" ~ "Settlements", 
    Legend == "Developed, High Intensity" ~ "Settlements", 
    Legend == "Barren Land" ~ "Bare Areas",
    Legend == "Deciduous Forest" ~ "Forest", 
    Legend == "Evergreen Forest" ~ "Forest", 
    Legend == "Mixed Forest" ~ "Forest", 
    Legend == "Shrub/Scrub" ~ "Grassland/Shrubland", 
    Legend == "Herbaceuous" ~ "Grassland/Shrubland", 
    Legend == "Hay/Pasture" ~ "Croplands", 
    Legend == "Woody Wetlands" ~ "Wetlands", 
    Legend == "Emergent Herbaceuous Wetlands" ~ "Wetlands",
  ))

print("Catchment landcover:")
lc <- veazie_lc %>% group_by(gres_class) %>% summarize(perc = sum(percent) %>% round(1)) 
print(lc)

# country
country <- "United States"
print(paste0("Country: ", country))

# coordinates
long <- -68.7011
lat <- 44.8322
print(paste0("Coordinates: ", lat, ",", long))

# reservoir area (km2)
res_area <- st_area(res) 
print(paste0("Reservoir Area (km2): ", res_area * 1e-6 ))

# mean depth (m)
mean_depth <- 4.42
print(paste0("Mean Depth (m): ", mean_depth))

# max depth (m)
max_depth <- 8.70
print(paste0("Max Depth (m): ", max_depth))

# littoral area (%) Output from g-res given from max and mean depth
lit_area <- NA
print(paste0("Littoral Area (%): ", lit_area))

# thermocline depth (m) Output from g-res
thermo_depth <- NA
print(paste0("Thermocline depth (m): ", thermo_depth))

# soil carbon
soil_carbon <- 5.11
print(paste0("Soil carbon under impounded area (kgC/m2): ", soil_carbon))

# reservoir wind
res_wind <- 5.31
print(paste0("Reservoir wind at 50 m: ", res_wind))

# water residence time (yrs) Output from g-res
wrt <- NA
print(paste0("Water residence time (yrs): ", wrt))

# annual discharge
q_annual <- 345
print(paste0("Annual Discharge (m3/s): ", q_annual))

# p concentration (ug/L) Output from Cronan et al. 2012
p_conc <- 3
print(paste0("P Concentration (ug/L): ", p_conc))

# global horizontal radiance
rad <- 3.47
print(paste0("Global horizontal radiance: ", rad))

# mean temperature per month
temp <- data.frame(Jan = -7.9, Feb = -6.7, Mar = -1.1, Apr = 5.4, May = 11.7, June = 17.0, July = 19.9, Aug = 18.8, Sept = 14.0, Oct = 8.4, Nov = 2.5, Dec = -4.8) %>% pivot_longer(cols = Jan:Dec, names_to = "Month", values_to = "Temperature")
print("Temperature per month (C)")
temp

# eventually it would be cool to build this out using a web feature service but for now we could use elevator to get 30 m elevation data or download 3DEP from the National Map

# res_slope <- terrain(elev, opt = "slope", unit = "degrees")
# sed_expo_area <- st_area(expo_sed)
# 
# # TODO add LC calculation
# perc_forest <- NULL
# perc_crop <- NULL
# 
# river_q <- NULL
```

```{r}
# library(soilDB)
# 
# # from Soil Grids not SSURGO or STATSGO
# 
# veazie_soil <- fetchSoilGrids(ws, depth_intervals = c("0-5"), variables = c("soc", "bdod"), grid = TRUE, progress = TRUE)
# 
# veazie_polyg <- ws %>% st_transform(crs(veazie_soil))
# 
# veazie_clip <- crop(veazie_soil, extent(veazie_polyg))
# veazie_mask <- mask(veazie_clip, veazie_polyg)
# 
# plot(veazie_mask)
# 
# names(veazie_mask)
# 
# veazie_c <- mean(values(veazie_mask[[7]], na.rm = T))
# veazie_bd <- mean(values(veazie_mask[[8]], na.rm = T))
# 
# # our estimates are definitely low and I suspect a unit conversion error but I'm not sure 
# veazie_c * 1e-1 * 1e-3 * veazie_bd * 1e-2 * 5 * 1e4
# 
# 
# # units are automatically converted but I think this is still wrong
# # https://search.r-project.org/CRAN/refmans/soilDB/html/fetchSoilGrids.html
# veazie_c * veazie_bd * 5 * 10
# 
# 100000 == 10^5
# 
# 1636*85.96*5*10^-5*10^-1*10^4
# 
# names(c_content_pared)
# # should I be concerned this is below the range
# range(c_content_pared$SOC_0_5, na.rm = TRUE)
# 
# #466
# 
# test_ws <- read_sf("../1-input-data/6-clow-data/test_ws/layers/globalwatershed.shp")
# 
# test_soil <- fetchSoilGrids(test_ws, depth_intervals = c("0-5"), variables = c("soc", "bdod"), grid = TRUE, progress = TRUE)
# 
# c_content %>% st_as_sf(coords = c("LON_DD", "LAT_DD"), crs = 4326) %>% dplyr::select(SOC_0_5) %>%  mapview()
```


# Before

## Reservoir surface emissions

**G-res tool**

https://131.datatrium.com/fmi/webd/G-res%20Tool?script=ChoiceWebPage&param=Grestool&homeurl=https://g-res.hydropower.org

**Required data**

Must know

- reservoir age
- P concentration (ug/L) or population in the catchment and category of wastewater treatment
- water residence time or mean depth
- reservoir area (km^2)
- reservoir climate zone
- maximum depth
- mean depth or volume

Can be derived from other inputs

- % littoral area
- water residence time

Can be calculated from Google Earth Engine script given a delineated catchment with 
https://code.earthengine.google.com/db5542fcc97469961bfd890123aa55df

- mean annual air temperature (C)
- reservoir surface soil carbon content (kgC m^-2)
- cumulative global horizontal radiance (kWh m^-2 d-1)
- catchment area (km2)
- annual runoff (mm/yr)
- catchment landcover (%)
- k factor (soil erodibility factor)
- mean basin slope

Can delineate catchments in US reservoirs using the StreamStats tool 
https://streamstats.usgs.gov/ss/ 

DONE
```{r surface emissions}
## output from g-res: emissions in g CO2 eq by pathway
res_co2_diff_eq <- 3
res_ch4_diff_eq <- 79 * .89
res_ch4_eb_eq <- 79 * 0.11
res_ch4_degas_eq <- 0

## without degassing
# calculate emissions in g CO2 eq without degassing
res_eq <- res_co2_diff_eq+res_ch4_diff_eq+res_ch4_eb_eq

# calculate emissions in g C without degassing
res_g <- (res_co2_diff_eq * (12.01/44.01)) + ((res_ch4_diff_eq+res_ch4_eb_eq) *(12.01/16.04) * (1/34))

# calculate standard deviation in g of CO2 eq without degassing
res_net_eq <- 64
res_net_2.5_eq <- 55
res_net_97.5_eq <- 75

res_eq_2.5 <- res_eq - max(c(res_net_97.5_eq-res_net_eq, res_net_eq - res_net_2.5_eq))
res_eq_97.5 <-  res_eq + max(c(res_net_97.5_eq-res_net_eq, res_net_eq - res_net_2.5_eq))

res_g_interval <- max(c(res_net_97.5_eq-res_net_eq, res_net_eq - res_net_2.5_eq))*(res_co2_diff_eq/res_eq)*(12.01/44.01) + max(c(res_net_97.5_eq-res_net_eq, res_net_eq - res_net_2.5_eq))*((res_ch4_diff_eq + res_ch4_degas_eq)/res_eq)*(1/34) * (12.01/16.04)

res_g_2.5 <- res_g - res_g_interval
res_g_97.5 <- res_g + res_g_interval

# got rid of the over-complicated combining of degassing an non-degassing fluxes
res_emissions_summary <- data.frame(period = c("before"), flux = c("reservoir emissions"), variation = NA, flux_unit = c("CO2-eq", "C"), annual = "yes", area = "reservoir", mean = c(res_eq, res_g), ci_lwr = c(res_eq_2.5, res_g_2.5), ci_upr = c(res_eq_97.5, res_g_97.5))
```

## Reservoir carbon burial TODO
```{r carbon burial}
# sedimentation model
clow_df <- read.csv("../1-input-data/6-clow-data/clow_all.csv") %>% mutate(sed_rate_m3_yr = sed_rate_mm_yr * 1e-3 * surface_area_m2, log_sed_rate_m3_yr = log10(sed_rate_m3_yr), log_surface_area_m2 = log10(surface_area_m2), crop_perc = Crop_Perc/100, forest_perc = Forest_Perc/100)

clow.mod <- lm(log_sed_rate_m3_yr ~ log_surface_area_m2 + slope_calc + forest_perc + crop_perc, data = clow_df)

# run predictors in the clow-predictors script 
# decided to use the same lc as input for g-res rather than EnviroAtlas delineation method because this also let's us get barren and it's weird to have different numbers here
veazie_sed_pred <- read.csv("../1-input-data/6-clow-data/veazie_sed_pred.csv") %>% 
  mutate(log_surface_area_m2 = log10(st_area(res) %>% as.numeric()), crop_perc = lc %>% filter(gres_class == "Croplands") %>% pull(perc)/100, forest_perc = lc %>% filter(gres_class == "Forest") %>% pull(perc)/100) %>% 
  dplyr::select(-Crop_Perc, -Forest_Perc, - dam) %>% 
  rename("slope_calc" = "slope")

# need to figure out how to use this prediction interval to sample sedimentation rates
predict(clow.mod, veazie_sed_pred, interval = "prediction")

# c content model
c_content <- read.csv("../1-input-data/6-clow-data/c_content.csv") %>% mutate(LAKE_ORIGIN_renamed = as.factor(LAKE_ORIGIN_renamed))

c_content_pared <- c_content %>% dplyr::select(sediment_C_pct, SOC_0_5, Wetlands90, log_Wetlands90, Barren30, KFactor, log_WaterBody_Area_m2, log_Lake_Area_m2_EPA, LAKE_ORIGIN_renamed) %>% mutate(wetland_si = log10(Wetlands90 + 1))

# I think this is the correct model 
c.mod <- lm(sediment_C_pct ~ SOC_0_5 + wetland_si + Barren30 + KFactor + log_Lake_Area_m2_EPA + LAKE_ORIGIN_renamed, data = c_content_pared)

# this is with kf not kw
veazie_c_pred <- read.csv("../1-input-data/6-clow-data/veazie_soil.csv") %>% rename("SOC_0_5" = "soc", "KFactor" = "kf") %>% dplyr::select(-kw) %>% mutate(wetland_si = log10(lc %>% filter(gres_class == "Wetlands") %>% pull(perc)/100 + 1), Barren30 = lc %>% filter(gres_class == "Bare Areas") %>% pull(perc)/100, log_Lake_Area_m2_EPA = log10(as.numeric(res_area)), LAKE_ORIGIN_renamed = "RESERVOIR")

predict(c.mod, veazie_c_pred, interval = "prediction")

```

```{r}
# spot checked that the log areas are correct but there are two different ones and who knows which they used

range(c_content$sediment_C_pct, na.rm = T)



c.mod <- lm(sediment_C_pct ~ SOC_0_5 + test_wetlands + Barren30 + KFactor + log_WaterBody_Area_m2 + wb_type, data = c_content %>% mutate(test_wetlands = log10(Wetlands90 + 1)))


options(scipen = 13)
summary(c.mod)

range(c_content$Wetlands90)

res_burial_summary <- data.frame(period = c("before"), flux = c("reservoir burial"), variation = NA, flux_unit = c("CO2-eq", "C"), annual = "yes", area = "reservoir", mean = c(-res_burial_co2eq, -res_burial), ci_lwr = c(-res_burial_co2eq_2.5, -res_burial_2.5), ci_upr = c(-res_burial_co2eq_97.5, -res_burial_97.5))
```

# Burp

## Exposed sediment CO2 emissions DONE
```{r}
# double check when you are fresh that these should be the same as the glines output
res_exp_sed_co2 <- read.csv("../2-output-data/exp_sed_co2.csv")

res_sed_co2_summary <- data.frame(period = c("burp"), flux = c("exposed sediment CO2"), variation = res_exp_sed_co2$y %>% as.character(), flux_unit = c(rep("CO2-eq", 6), rep("C", 6)), annual = "no", area = "exposed sediment area", mean = c(res_exp_sed_co2$mean_co2, res_exp_sed_co2$mean_c), ci_lwr = c(res_exp_sed_co2$ci_lower_co2, res_exp_sed_co2$ci_lower_c), ci_upr = c(res_exp_sed_co2$ci_upper_co2, res_exp_sed_co2$ci_upper_c))
```

## Exposed sediment CH4 emissions DONE
```{r}
thresh_0.01 <- read.csv("../2-output-data/exp_sed_ch4.csv")

pore_co2_eq <- 8102.2 * 1e-3 * (16.04/12.01)* 34
pore_c <- 8102.2 * 1e-3

res_sed_ch4_summary <- data.frame(period = c("burp"), flux = c("exposed sediment CH4"), variation = c(rep(paste0("no_pore_", thresh_0.01$y), 2), rep(paste0("pore_", thresh_0.01$y), 2)), flux_unit = rep(c(rep("CO2-eq", 6), rep("C", 6)), 2), annual = "no", area = "exposed sediment area", mean = c(thresh_0.01$mean_co2,  thresh_0.01$mean_c, thresh_0.01$mean_co2 + pore_co2_eq, thresh_0.01$mean_c + pore_c), ci_lwr = c(thresh_0.01$lwr_co2,  thresh_0.01$lwr_c, thresh_0.01$lwr_co2 + pore_co2_eq, thresh_0.01$lwr_c + pore_c), ci_upr = c(thresh_0.01$upr_co2, thresh_0.01$upr_c, thresh_0.01$upr_co2 + pore_co2_eq, thresh_0.01$upr_c + pore_c))
```

## Drawdown ebullition DONE
```{r}
res_ch4_eb_eq_2.5 <- res_ch4_eb_eq - max((res_net_eq - res_net_2.5_eq), (res_net_97.5_eq - res_net_eq))
res_ch4_eb_eq_97.5 <- res_ch4_eb_eq + max((res_net_eq - res_net_2.5_eq), (res_net_97.5_eq - res_net_eq))

drawdown_eb_prop <- c(0, 5, 10, 25, 100)
round(res_ch4_eb_eq * drawdown_eb_prop,0)
round(res_ch4_eb_eq_2.5 * drawdown_eb_prop,0)
round(res_ch4_eb_eq_97.5 * drawdown_eb_prop,0)

res_ch4_eb_g <- res_ch4_eb_eq * (1/34) * (12.01/16.04)
res_ch4_eb_g_2.5 <- res_ch4_eb_eq_2.5 * (1/34) * (12.01/16.04)
res_ch4_eb_g_97.5 <- res_ch4_eb_eq_97.5 * (1/34) * (12.01/16.04)

round(res_ch4_eb_g * drawdown_eb_prop,1)
round(res_ch4_eb_g_2.5 * drawdown_eb_prop,1)
round(res_ch4_eb_g_97.5 * drawdown_eb_prop,1)

res_drawdown_eb_summary <- data.frame(period = c("burp"), flux = c("drawdown ebullition"), variation = drawdown_eb_prop %>% as.character(), flux_unit = c(rep("CO2-eq",5), rep("C", 5)), annual = "no", area = "reservoir", mean = c(res_ch4_eb_eq * drawdown_eb_prop, res_ch4_eb_g * drawdown_eb_prop) , ci_lwr = c(res_ch4_eb_eq_2.5 * drawdown_eb_prop, res_ch4_eb_g_2.5 * drawdown_eb_prop) , ci_upr = c(res_ch4_eb_eq_97.5 * drawdown_eb_prop, res_ch4_eb_g_97.5 * drawdown_eb_prop))
```

## Eroded sediment emissions TODO
```{r}
# Collins et al. write that a combination of the CR Environmental Report and post-removal survey suggest that the combined volume of mobilized sediment from Veazie and Great Works dam was 15,000 m3 or smaller
# Fit a uniform distribution between 0 and 15,000 m3
res_erode_g_max <- 15000
res_erode_g_min <- 0

# parameters for g distribution of % C
b = (c_content_sd)^2/c_content
a = (c_content)^2/(c_content_sd)^2

# assumming between 0 and 100% oxidized
min_ox_trans = 0 
max_ox_trans = 1

eroded_vec <- c()
for(i in 1:10000){
  temporary <- (runif(1, min = res_erode_g_min, max = res_erode_g_max) * rgamma(1, shape=a, scale = b) * runif(1, min = min_ox_trans, max = max_ox_trans))/(res_area * 1e6)
  eroded_vec <- c(eroded_vec, temporary)
}

res_eroded_97.5 <- sort(eroded_vec)[9750]
res_eroded <- sort(eroded_vec)[5000]
res_eroded_2.5 <- sort(eroded_vec)[250]

res_eroded_co2eq_97.5 <- res_eroded_97.5 * (44.01/12.01)
res_eroded_co2eq <- res_eroded * (44.01/12.01)
res_eroded_co2eq_2.5 <- res_eroded_2.5 * (44.01/12.01)

# these are per m2
print(paste0("eroded sediment emissions in g C m-2: ", round(res_eroded,3), " (", round(res_eroded_2.5, 3), " - ", round(res_eroded_97.5, 3), ")")) 

print(paste0("eroded sediment emissions in g CO2eq m-2: ", round(res_eroded_co2eq,3), " (", round(res_eroded_co2eq_2.5, 3), " - ", round(res_eroded_co2eq_97.5, 3), ")")) 

res_erode_summary <- data.frame(period = c("burp"), flux = c("eroded sediment emissions"), variation = NA, flux_unit = c("CO2-eq", "C"), annual = "no", area = "reservoir", mean = c(res_eroded_co2eq, res_eroded), ci_lwr = c(res_eroded_co2eq_2.5, res_eroded_2.5), ci_upr = c(res_eroded_co2eq_97.5, res_eroded_97.5))
```


# Build

## River surface CO2 emissions DONE
```{r}
# Nate suggested writing a function that takes parameters of a gamma distribution as arguments and calculates those parameters that I have. Use optim or mle2 to optimize estimates to be as close to empirical as possible

# Discharge is in the >100 m3/s bin
# Values are reported as g C m-2 d-

# Function to calculate the sum of squared differences between observed and target quantiles

gamma_params_diff <- function(params, quantiles_target) {
  shape <- params[1]
  scale <- params[2]
  
  quantiles_observed <- qgamma(c(0.25, 0.5, 0.75), shape = shape, scale = scale)
  
  sum((quantiles_observed - quantiles_target)^2)
}

# Target quantiles
quantiles_target <- c(1.5, 2.16, 4.78)

# Initial guess for shape and scale parameters
initial_params <- c(shape = 1, scale = 1)

# Optimize the parameters
optimized_params <- optim(par = initial_params, fn = gamma_params_diff, quantiles_target = quantiles_target)

# Extract optimized parameters
shape_optimized <- optimized_params$par[1]
scale_optimized <- optimized_params$par[2]

# Check the quantiles with the optimized parameters
quantiles_optimized <- qgamma(c(0.25, 0.5, 0.75), shape = shape_optimized, scale = scale_optimized)
cat("Optimized Quantiles:", quantiles_optimized, "\n")

x <- seq(0, 10, by = 0.1)
y <- dgamma(x, shape = shape_optimized, scale = scale_optimized)

plot(x, y)

riv_co2_vec <- sort(rgamma(10000, shape = shape_optimized, scale = scale_optimized)) *365

riv_co2_97.5_gC <- riv_co2_vec[9750]
riv_co2_gC <- riv_co2_vec[5000]
riv_co2_2.5_gC <- riv_co2_vec[250]

riv_co2_2.5_gC
riv_co2_gC
riv_co2_97.5_gC

riv_co2_vec_gco2 <- riv_co2_vec * (44.01/12.01)

riv_co2_97.5_gco2 <- riv_co2_vec_gco2[9750]
riv_co2_gco2 <- riv_co2_vec_gco2[5000]
riv_co2_2.5_gco2 <- riv_co2_vec_gco2[250]

riv_co2_97.5_gco2
riv_co2_gco2
riv_co2_2.5_gco2

riv_co2_summary <- data.frame(period = c("build"), flux = c("river CO2 emissions"), variation = NA, flux_unit = c("CO2-eq", "C"), annual = "yes", area = "river", mean = c(riv_co2_gco2, riv_co2_gC), ci_lwr = c(riv_co2_2.5_gco2, riv_co2_2.5_gC), ci_upr = c(riv_co2_97.5_gco2, riv_co2_97.5_gC))
```

## River surface CH4 emissions DONE these seem, really high
```{r}
ch4_conc <- read.csv("../1-input-data/5-ch4-fluxes/GRiMe_concentrations_v2.csv") 
ch4_flux <- read.csv("../1-input-data/5-ch4-fluxes/GRiMe_fluxes_v2.csv")
ch4_flux_sites <- read.csv("../1-input-data/5-ch4-fluxes/GRiMe_sites_v2.csv")
ch4_source <- read.csv("../1-input-data/5-ch4-fluxes/GRiMe_sources_v2.csv")

usa <- ne_countries(country = "united states of america")

ch4_flux_sf <- ch4_flux %>% 
  left_join(ch4_flux_sites, by = c("Source_ID", "Site_ID", "Site_Name")) %>% 
  left_join(ch4_conc %>% 
              filter(FluxYesNo == "Yes") %>% 
              rename("Flux_Name" = "Conc_Name"), 
            by = c("Source_ID", "Site_ID", "Site_Name", "Flux_Name")) %>% 
  filter(is.na(Longitude)==FALSE, is.na(Channel_type)==TRUE) %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326) 

# filter sites in us
ch4_flux_sites_usa <- st_intersection(usa, ch4_flux_sf) %>% dplyr::select(Source_ID:geometry)
mapview(ch4_flux_sites_usa)

# Does it make sense to bin by discharge just because Hotckiss et al. 2015 did if discharge is not explanatory?
ch4_flux_sites_usa %>% ggplot(aes(Q, Diffusive_CH4_Flux_Mean))+geom_point()
ch4_flux_sites_usa %>% filter(Q < 500 & Diffusive_CH4_Flux_Mean<45) %>% ggplot(aes(Q, Diffusive_CH4_Flux_Mean))+geom_point()


# TODO: figure out how many obs are in each bin, sample appropriate bin report number, decide what to do RE ebullition, and compare CO2 values to the ones derived from Hotchkiss et al. 2015

ch4_flux_sites_usa <- ch4_flux_sites_usa %>% mutate(q_bin = case_when(
  (Q>=0 & Q<0.01) ~ "0-0.01",
  (Q>0.01 & Q<0.1) ~ "0.01-0.1",
  (Q>0.1 & Q<1) ~ "0.1-1",
  (Q>1 & Q<10) ~ "1-10",
  (Q>10 & Q<100) ~ "10-100",
  (Q > 100) ~ ">100"
)) 

# summary of diffusive CH4 by bin
ch4_flux_sites_usa %>% filter(is.na(Diffusive_CH4_Flux_Mean)==FALSE) %>% group_by(q_bin) %>% summarize(count = n())

# no measurements of ebullition in the Q range so assuming it is zero. This is another caveat we will need to address
ch4_flux_sites_usa %>% filter(is.na(Eb_CH4_Flux_Mean)==FALSE) %>% group_by(q_bin) %>% summarize(count = n())


ch4_flux_vec <- sort(sample(ch4_flux_sites_usa %>% filter(q_bin== ">100") %>% pull(Diffusive_CH4_Flux_Mean), 10000, replace = TRUE))
ch4_flux_vec_gCH4 <- ch4_flux_vec * 365 * 1e-3 * 16.04
ch4_flux_vec_gCO2eq <- ch4_flux_vec_gCH4 * 34

# could we just calculate the quantiles
# quantile(ch4_flux_sites_usa %>% filter(q_bin== "10-100") %>% pull(Diffusive_CH4_Flux_Mean) * 365 * 1e-3 * 16.04, c(0.025, 0.5, 0.975))

riv_ch4_flux_97.5_gC <- ch4_flux_vec_gCH4[9750] * (12.01 / 16.04)
riv_ch4_flux_gC <- ch4_flux_vec_gCH4[5000] * (12.01 / 16.04)
riv_ch4_flux_2.5_gC <- ch4_flux_vec_gCH4[250] * (12.01 / 16.04)

riv_ch4_flux_97.5_gC
riv_ch4_flux_gC
riv_ch4_flux_2.5_gC

riv_ch4_flux_97.5_gCO2eq <- ch4_flux_vec_gCO2eq[9750]
riv_ch4_flux_gCO2eq <- ch4_flux_vec_gCO2eq[5000]
riv_ch4_flux_2.5_gCO2eq <- ch4_flux_vec_gCO2eq[250]

riv_ch4_flux_97.5_gCO2eq
riv_ch4_flux_gCO2eq
riv_ch4_flux_2.5_gCO2eq

# CO2 flux
co2_vec <- sort(sample(ch4_flux_sites_usa %>% filter(is.na(CO2_Flux_Mean)==FALSE)  %>% filter(q_bin== "10-100") %>% pull(CO2_Flux_Mean), 10000, replace = TRUE))* 365 * 1e-3 * 44.01
co2_vec[250]
co2_vec[5000]
co2_vec[9750]

riv_ch4_summary <- data.frame(period = c("build"), flux = c("river CH4 emissions"), variation = NA, flux_unit = c("CO2-eq", "C"), annual = "yes", area = "river", mean = c(riv_ch4_flux_gCO2eq, riv_ch4_flux_gC), ci_lwr = c(riv_ch4_flux_2.5_gCO2eq, riv_ch4_flux_2.5_gC), ci_upr = c(riv_ch4_flux_97.5_gCO2eq, riv_ch4_flux_97.5_gC))
```

## NEE DONE
```{r}
f_age <- function(a, b, e, age){
  return(e + a * (1-exp(b*age)))
}

# changed inputs from a = 587.74, b = -0.19, e = -324.54 to a = 263.2, b = -0.19, e = 0 to make it reach the same asymptote but start at 0
# it doesn't seem to matter what the reference level is or what the f_age parameters are (in terms of the two options above, we get the same relationship)
df <- read.csv("../1-input-data/8-besnard-data/besnard_2018.csv") %>% mutate(age_func = f_age(263.2, -0.19, 0, Age)) %>% mutate(nutrient = as.factor(`NA.`))

# estimating NEP assuming that it is zero before the model predicts positive values
nep_mod <- lm(NEP ~ MAT + Age + age_func + nutrient, data = df)
summary(nep_mod)

nep_new <- data.frame(MAT = mean(temp$Temperature), Age = seq(0, 100, by = 1), age_func = f_age(263.2, -0.19, 0, seq(0, 100, by = 1)), nutrient =as.factor("L"))
nep_pred <- predict(nep_mod, nep_new, interval = "prediction") %>% as.data.frame()

# becomes 0 a little before 5.7 years in so going to call it 6

nep_df <- data.frame(years = double(), mean_gC = double(), lwr_gC = double(), upr_gC = double())

year_vec <- c(25,50,100)
for(i in 1:length(year_vec)){
  yr <- year_vec[i]
  mean_gC <- integrate(function(x) as.data.frame(predict(nep_mod, newdata = data.frame(MAT = mean(temp$Temperature), Age = x, age_func = f_age(263.2, -0.19, 0, x), nutrient =as.factor("L")), interval = "prediction", level = 0.95))$fit, lower = 6, upper = yr)$value
  lwr_gC <- -(integrate(function(x) -(as.data.frame(predict(nep_mod, newdata = data.frame(MAT = mean(temp$Temperature), Age = x, age_func = f_age(263.2, -0.19, 0, x), nutrient =as.factor("L")), interval = "prediction", level = 0.95))$lwr), lower = 6, upper = yr))$value
  upr_gC <- integrate(function(x) as.data.frame(predict(nep_mod, newdata = data.frame(MAT = mean(temp$Temperature), Age = x, age_func = f_age(263.2, -0.19, 0, x), nutrient =as.factor("L")), interval = "prediction", level = 0.95))$upr, lower = 6, upper = yr)$value
  
  temporary <- data.frame(years = yr, mean_gC = -mean_gC, lwr_gC = -lwr_gC, upr_gC = -upr_gC)
  
  nep_df <- nep_df %>% bind_rows(temporary)
}

# integrated carbon
nep_df <- nep_df %>% mutate(mean_gCO2 = mean_gC * (44.01/12.01), lwr_gCO2 = lwr_gC * (44.01/12.01), upr_gCO2 = upr_gC * (44.01/12.01))

# average annual rates
nep_df_rates <- nep_df %>% mutate(across(mean_gC:upr_gCO2, ~ .x / years))

nep_df_rates

res_nep_summary <- data.frame(period = c("build"), flux = c("NEP"), variation = nep_df_rates$years %>% as.character(), flux_unit = c(rep("CO2-eq", 3), rep("C", 3)), annual = "yes", area = "exposed sediment area", mean = c(nep_df_rates$mean_gCO2, nep_df_rates$mean_gC), ci_lwr = c(nep_df_rates$lwr_gCO2, nep_df_rates$lwr_gC), ci_upr = c(nep_df_rates$upr_gCO2, nep_df_rates$upr_gC))

res_nep_summary
```

## Soil CH4 emissions DONE
## Priority for math checking
```{r}
soil <- read.csv("../1-input-data/7-gattica-data/MethaneSoilForest.csv") %>% mutate(MAP_sq = MAP^2, MAT_sq = MAT^2) %>% filter(complete.cases(.))

soil_scale <- soil %>% mutate(across(c(MAP:soilpH, MAP_sq, MAT_sq), scale))

soil_mod <- lmer(Methane ~ MAP*MAT + soilpH + SOC + BulkDens + MAT_sq + SoilNit + MAP_sq + SoilSand + (1 + MAP*MAT + soilpH + SOC + BulkDens + MAT_sq + SoilNit + MAP_sq + SoilSand| Biome), data = soil_scale)

# I am not getting the same AICc
summary(soil_mod)
AICc(soil_mod, REML = FALSE)
AICc(soil_mod, REML = TRUE)

# need to get the predictors for sites
# MAP, MAT, soilpH, BulkDens, SoilNit, Soil Sand

veazie_soil <- read.csv("../1-input-data/7-gattica-data/veazie_soil.csv")

# going to have to manually scale                                                         
soil_data_new <- data.frame(MAP = (veazie_soil$veazie_map- mean(soil$MAP))/sd(soil$MAP),
                            MAT = (veazie_soil$veazie_mat - mean(soil$MAT))/sd(soil$MAT), 
                            soilpH = (veazie_soil$veazie_ph - mean(soil$soilpH))/sd(soil$soilpH), 
                            SOC = (veazie_soil$veazie_oc - mean(soil$SOC))/sd(soil$SOC), 
                            BulkDens = (veazie_soil$veazie_bd - mean(soil$BulkDens))/sd(soil$BulkDens), 
                            SoilNit = (veazie_soil$veazie_tn - mean(soil$SoilNit))/sd(soil$SoilNit), 
                            SoilSand = (veazie_soil$veazie_sand - mean(soil$SoilSand))/sd(soil$SoilSand), 
                            MAT_sq = (veazie_soil$veazie_mat^2 - mean(soil$MAT_sq))/sd(soil$MAT_sq), 
                            MAP_sq = (veazie_soil$veazie_map^2 - mean(soil$MAP_sq))/sd(soil$MAP_sq), 
                            Biome = "Temperate")

predict(soil_mod, soil_data_new)

library(merTools)

# returns values in kg Ch4 /ha/yr
soil_ch4_out <- predictInterval(soil_mod, soil_data_new, n.sims = 10000, level = 0.95) %>% as.data.frame()

soil_ch4_out <- soil_ch4_out %>% mutate(soil_c = fit * 0.1 * (12.01/16.04), 
                        soil_c_2.5 = lwr * 0.1 * (12.01/16.04), 
                        soil_c_97.5 = upr * 0.1 * (12.01/16.04), 
                        soil_co2_eq = fit * 0.1 * 34, 
                        soil_co2_eq_2.5 = lwr * 0.1 * 34, 
                        soil_co2_eq_97.5 = upr * 0.1 * 34)  
  
soil_ch4_out
# pivot_longer(cols = c(fit:soil_co2_eq_97.5), 
  #              values_to = "value",
  #              names_to = "name")

res_soil_ch4_summary <- data.frame(period = c("build"), flux = c("soil CH4 emissions"), variation = NA, flux_unit = c("CO2-eq", "C"), annual = "yes", area = "exposed sediment area", mean = c(soil_ch4_out$soil_co2_eq, soil_ch4_out$soil_c), ci_lwr = c(soil_ch4_out$soil_co2_eq_2.5, soil_ch4_out$soil_c_2.5), ci_upr = c(soil_ch4_out$soil_co2_eq_97.5, soil_ch4_out$soil_c_97.5))
```


## Tree CH4 emissions TODO
```{r}
# TODO calc slope, aspect, and elevation as inputs for fvs in R not arcpro
terrace <- st_read("../1-input-data/1-geospatial/post-removal/lake-mills_terrace.shp")
valley_wall <- st_read("../1-input-data/1-geospatial/post-removal/lake-mills_valley-wall.shp")

# dtm <- raster("../1-input-data/1-geospatial/post-removal/lake-mills_post-removal_dtm.tif")

terrace_slope <- read.csv("../1-input-data/1-geospatial/post-removal/lake-mills_terrace_slope.csv") %>% dplyr::select(stand_id, MEAN) %>% rename("slope" = "MEAN")
terrace_elevation <- read.csv("../1-input-data/1-geospatial/post-removal/lake-mills_terrace_elevation.csv") %>% dplyr::select(stand_id, MEAN) %>% rename("elevation" = "MEAN")
terrace_aspect <- read.csv("../1-input-data/1-geospatial/post-removal/lake-mills_terrace_aspect.csv") %>% dplyr::select(stand_id, MEAN) %>% rename("aspect" = "MEAN")

# coordinates for FVS input
terrace_coords <- terrace %>% st_transform(4326) %>% st_centroid() %>% st_coordinates() %>% as.data.frame()
terrace_coords$stand_id <- terrace_slope$stand_id

# terrace inputs
terrace_inputs <- terrace_slope %>% left_join(terrace_elevation, by = "stand_id") %>% left_join(terrace_aspect, by = "stand_id") %>% left_join(terrace_coords, by = "stand_id")

valley_wall_slope <- read.csv("../1-input-data/1-geospatial/post-removal/lake-mills_valley-wall_slope.csv") %>% dplyr::select(stand_id, MEAN) %>% rename("slope" = "MEAN")
valley_wall_aspect <- read.csv("../1-input-data/1-geospatial/post-removal/lake-mills_valley-wall_aspect.csv") %>% dplyr::select(stand_id, MEAN) %>% rename("aspect" = "MEAN")
valley_wall_elevation <- read.csv("../1-input-data/1-geospatial/post-removal/lake-mills_valley-wall_elevation.csv") %>% dplyr::select(stand_id, MEAN) %>% rename("elevation" = "MEAN")

valley_wall_coords <- valley_wall %>% st_transform(4326) %>% st_centroid() %>% st_coordinates() %>% as.data.frame()
valley_wall_coords$stand_id <- valley_wall_slope$stand_id

valley_wall_inputs <- valley_wall_slope %>% left_join(valley_wall_elevation, by = "stand_id") %>% left_join(valley_wall_aspect, by = "stand_id")%>% left_join(valley_wall_coords, by = "stand_id")

# write.csv(valley_wall_inputs, "../1-input-data/2-fvs/lake-mills_valley-wall.csv")

trees_k09 <- read.csv("../1-input-data/2-fvs/lake-mills-output_k09.csv")
trees_k15 <- read.csv("../1-input-data/2-fvs/lake-mills-output_k15.csv")

# DBH in inches Ht in ft convert to m2
tree_surface_area_k09 <- trees_k09 %>% 
  mutate(surface_area = (DBH/12)*pi *EstHt *0.092903, surface_area_3ft = (DBH/12)*pi * 3 *0.092903) %>% 
  group_by(StandID, Year) %>% summarize(total_surface_area = sum(surface_area), total_surface_area_3ft = sum(surface_area_3ft)) %>% mutate(habitat = if_else(str_detect(StandID, "lmt"), "wetland", "upland"))

tree_surface_area_k15 <- trees_k15 %>% 
  mutate(surface_area = (DBH/12)*pi *EstHt * 0.092903, surface_area_3ft = (DBH/12)*pi * 3* 0.092903) %>% 
  group_by(StandID, Year) %>% summarize(total_surface_area = sum(surface_area), total_surface_area_3ft = sum(surface_area_3ft))%>% mutate(habitat = if_else(str_detect(StandID, "lmt"), "wetland", "upland"))

# ug CH4/m2/hr
upland_tree_ch4 <- 68.8 * 8760 *1e-6
upland_tree_ch4_sd <- 53.6 * 8760 *1e-6

wetland_tree_ch4 <- 567.9 * 8760 *1e-6
wetland_tree_ch4_sd <- 523.5 * 8760 *1e-6

tree_plots <- tree_surface_area_k09 %>% dplyr::select(StandID, habitat) %>% distinct()
tree_plots$area <- c(st_area(terrace) %>% as.numeric(), st_area(valley_wall) %>% as.numeric())
tree_plots

# report area-weighted average rate 
# tree_ch4_df <- NULL
# 
# for(i in 1:10000){
#   tree_plots$ch4_rate <- c(rnorm(7, wetland_tree_ch4, wetland_tree_ch4_sd), rnorm(3, upland_tree_ch4, upland_tree_ch4_sd))
#   
#   tree_sa <- sample(c(1, 2), size = 1, replace = TRUE)
#   
#   if(tree_sa == 1){
#     temporary <- tree_surface_area_k09 %>% 
#       left_join(tree_plots, by = c("StandID", "habitat")) %>% 
#       mutate(ch4_all = total_surface_area * ch4_rate, 
#              ch4_3ft = total_surface_area_3ft * ch4_rate) %>% 
#       group_by(StandID) %>% 
#       summarize(ch4_all_avg = mean(ch4_all), ch4_3ft_avg = mean(ch4_3ft)) %>% 
#       ungroup() %>% 
#       left_join(tree_plots %>% dplyr::select(-ch4_rate), by = c("StandID")) %>% 
#       mutate(ch4_all_rate = ch4_all_avg/area, ch4_3ft_rate = ch4_3ft_avg/area) %>% 
#       summarize(mean_tree_ch4_all = mean(ch4_all_rate), 
#                 mean_tree_ch4_3ft = mean(ch4_3ft_rate), 
#                 mean_tree_ch4_all_co2 = mean_tree_ch4_all * 34, 
#                 mean_tree_ch4_3ft_co2 = mean_tree_ch4_3ft * 34)
#       
#   }
#   
#   if(tree_sa == 2){
#     temporary <- tree_surface_area_k15 %>% 
#       left_join(tree_plots, by = c("StandID", "habitat")) %>% 
#       mutate(ch4_all = total_surface_area * ch4_rate, 
#              ch4_3ft = total_surface_area_3ft * ch4_rate) %>% 
#       group_by(StandID) %>% 
#       summarize(ch4_all_avg = mean(ch4_all), ch4_3ft_avg = mean(ch4_3ft)) %>% 
#       ungroup() %>% 
#       left_join(tree_plots %>% dplyr::select(-ch4_rate), by = c("StandID")) %>% 
#       mutate(ch4_all_rate = ch4_all_avg/area, ch4_3ft_rate = ch4_3ft_avg/area) %>% 
#       summarize(mean_tree_ch4_all = mean(ch4_all_rate), 
#                 mean_tree_ch4_3ft = mean(ch4_3ft_rate), 
#                 mean_tree_ch4_all_co2 = mean_tree_ch4_all * 34, 
#                 mean_tree_ch4_3ft_co2 = mean_tree_ch4_3ft * 34)
#     
#   }
#   
#   tree_ch4_df <- tree_ch4_df %>% bind_rows(temporary)
# }
# 
# 
# write.csv(tree_ch4_df, "../2-output-data/glines_tree_ch4_emissions.csv")

tree_ch4_df <- read.csv("../2-output-data/glines_tree_ch4_emissions.csv")

ch4_all <- tree_ch4_df$mean_tree_ch4_all %>% sort()
ch4_3ft <- tree_ch4_df$mean_tree_ch4_3ft %>% sort()

ch4_all_co2 <- tree_ch4_df$mean_tree_ch4_all_co2 %>% sort()
ch4_3ft_co2 <- tree_ch4_df$mean_tree_ch4_3ft_co2 %>% sort()

tree_ch4_97.5 <- ch4_all[9750] * (12.01/16.04)
tree_ch4 <- ch4_all[5000] * (12.01/16.04)
tree_ch4_2.5 <- ch4_all[250] * (12.01/16.04)

tree_ch4_3ft_97.5 <- ch4_3ft[9750]* (12.01/16.04)
tree_ch4_3ft <- ch4_3ft[5000]* (12.01/16.04)
tree_ch4_3ft_2.5 <- ch4_3ft[250]* (12.01/16.04)

tree_ch4_co2_97.5 <- ch4_all_co2[9750]
tree_ch4_co2 <- ch4_all_co2[5000]
tree_ch4_co2_2.5 <- ch4_all_co2[250]

tree_ch4_3ft_co2_97.5 <- ch4_3ft_co2[9750]
tree_ch4_3ft_co2 <- ch4_3ft_co2[5000]
tree_ch4_3ft_co2_2.5 <- ch4_3ft_co2[250]


tree_ch4
tree_ch4_2.5
tree_ch4_97.5

tree_ch4_co2
tree_ch4_co2_2.5
tree_ch4_co2_97.5

tree_ch4_3ft
tree_ch4_3ft_2.5
tree_ch4_3ft_97.5

tree_ch4_3ft_co2
tree_ch4_3ft_co2_2.5
tree_ch4_3ft_co2_97.5

res_tree_ch4_summary <- data.frame(period = c("build"), 
                                   flux = c("tree CH4 emissions"), 
                                   variation = c("offset by canopy", "offset by canopy", "whole trunk", "whole trunk", "three feet", "three feet"), 
                                   flux_unit = rep(c("CO2-eq", "C"), 3), 
                                   annual = "yes", 
                                   area = "exposed sediment area", 
                                   mean = c(0, 0, tree_ch4_co2, tree_ch4, tree_ch4_3ft_co2, tree_ch4_3ft), 
                                   ci_lwr = c(0, 0,tree_ch4_co2_2.5,  tree_ch4_2.5, tree_ch4_3ft_co2_2.5, tree_ch4_3ft_2.5), 
                                   ci_upr = c(0, 0, tree_ch4_co2_97.5, tree_ch4_97.5, tree_ch4_3ft_co2_97.5, tree_ch4_3ft_97.5))

res_tree_ch4_summary
```

# Summary TODO
```{r}
# The terrace delineation was done using 2014 NAIP data. The flow is now more concentrated in a smaller area meaning there is a greater area of exposed sediments. I don't think this matters because the FVS model is not that sensitive to small changes in aspect, elevation, and slope, so I am moving forward with 2021 delineations for sediment and river area
glines_riv <- read_sf("../1-input-data/1-geospatial/post-removal/elwha-river_glines.shp")
names(glines_riv)

glines_riv_area <- (glines_riv %>% filter(Type == "main") %>% st_area()) - (glines_riv %>% filter(Type != "main") %>% st_area() %>% sum()) 

area_df <- data.frame(area = c("reservoir", "exposed sediment area", "river"), 
                      area_m2 = c(res_area * 1e6, (res_area * 1e6)- as.numeric(glines_riv_area), as.numeric(glines_riv_area)))

# negative values are for uptake or burial positive values are for emissions
# how to show the multiple scenarios e.g., drawdown ebullition (only the most extreme or is there a way to layer on the plot)

summary_df <- bind_rows(res_emissions_summary, res_burial_summary, res_sed_co2_summary, res_sed_ch4_summary, res_drawdown_eb_summary, res_erode_summary, riv_co2_summary, riv_ch4_summary, res_nep_summary, res_soil_ch4_summary, res_tree_ch4_summary) %>% left_join(area_df, by = "area") %>% mutate(mean_area = mean * area_m2, ci_lwr_area = ci_lwr * area_m2, ci_upr_area = ci_upr * area_m2) %>% mutate(flux_fct = fct_relevel(flux, c("reservoir emissions", "reservoir burial", "exposed sediment CO2", "exposed sediment CH4", "drawdown ebullition", "eroded sediment emissions", "river CO2 emissions", "river CH4 emissions", "NEP", "soil CH4 emissions", "tree CH4 emissions")))

best_case_df <- bind_rows(res_emissions_summary, 
                          res_burial_summary, 
                          res_sed_co2_summary %>% filter(variation == "0.0833333333333333"),
                          res_sed_ch4_summary %>% filter(variation == "no_pore_0.0833333333333333"), 
                          res_drawdown_eb_summary %>% filter(variation == "0"), 
                          res_erode_summary, 
                          riv_co2_summary, 
                          riv_ch4_summary, 
                          res_nep_summary %>% filter(variation == "100"), 
                          res_soil_ch4_summary,
                          res_tree_ch4_summary %>% filter(variation == "offset by canopy")) %>% left_join(area_df, by = "area") %>% mutate(mean_area = mean * area_m2, ci_lwr_area = ci_lwr * area_m2, ci_upr_area = ci_upr * area_m2) %>% mutate(flux_fct = fct_relevel(flux, c("reservoir emissions", "reservoir burial", "exposed sediment CO2", "exposed sediment CH4", "drawdown ebullition", "eroded sediment emissions", "river CO2 emissions", "river CH4 emissions", "NEP", "soil CH4 emissions", "tree CH4 emissions")))

worst_case_df <- bind_rows(res_emissions_summary, 
                          res_burial_summary, 
                          res_sed_co2_summary %>% filter(variation == "100"),
                          res_sed_ch4_summary %>% filter(variation == "pore_100"), 
                          res_drawdown_eb_summary %>% filter(variation == "100"), 
                          res_erode_summary, 
                          riv_co2_summary, 
                          riv_ch4_summary, 
                          res_nep_summary %>% filter(variation == "100"), 
                          res_soil_ch4_summary,
                          res_tree_ch4_summary %>% filter(variation == "whole trunk")) %>% left_join(area_df, by = "area") %>% mutate(mean_area = mean * area_m2, ci_lwr_area = ci_lwr * area_m2, ci_upr_area = ci_upr * area_m2) %>% mutate(flux_fct = fct_relevel(flux, c("reservoir emissions", "reservoir burial", "exposed sediment CO2", "exposed sediment CH4", "drawdown ebullition", "eroded sediment emissions", "river CO2 emissions", "river CH4 emissions", "NEP", "soil CH4 emissions", "tree CH4 emissions")))


middle_case_df <-bind_rows(res_emissions_summary, 
                          res_burial_summary, 
                          res_sed_co2_summary %>% filter(variation == "10"),
                          res_sed_ch4_summary %>% filter(variation == "pore_10"), 
                          res_drawdown_eb_summary %>% filter(variation == "10"), 
                          res_erode_summary, 
                          riv_co2_summary, 
                          riv_ch4_summary, 
                          res_nep_summary %>% filter(variation == "100"), 
                          res_soil_ch4_summary,
                          res_tree_ch4_summary %>% filter(variation == "three feet")) %>% left_join(area_df, by = "area") %>% mutate(mean_area = mean * area_m2, ci_lwr_area = ci_lwr * area_m2, ci_upr_area = ci_upr * area_m2) %>% mutate(flux_fct = fct_relevel(flux, c("reservoir emissions", "reservoir burial", "exposed sediment CO2", "exposed sediment CH4", "drawdown ebullition", "eroded sediment emissions", "river CO2 emissions", "river CH4 emissions", "NEP", "soil CH4 emissions", "tree CH4 emissions")))
  

unique(res_tree_ch4_summary$variation)

png("../2-output-data/co2eq_draft1.png", width = 8, height = 5, units = "in", res = 300)
summary_df %>% filter(flux_unit == "CO2-eq") %>% ggplot(aes(flux_fct, mean))+geom_col(aes(fill = period)) + geom_errorbar(aes(x = flux_fct, ymin = ci_lwr, ymax = ci_upr)) + theme(axis.text.x = element_text(angle = 90)) + labs(y = "g CO2-eq/m2/yr", x = "")
dev.off()
```

```{r}
# colors: "#86BBD8", "#588157", "#848484"
png("../2-output-data/co2eq_draft2.png", width = 9, height = 5, units = "in", res = 300)
best_case_df %>% 
  filter(flux_unit == "CO2-eq", flux_fct %in% c("reservoir emissions", "reservoir burial", "soil CH4 emissions", "NEP", "tree CH4 emissions", "river CH4 emissions", "river CO2 emissions")) %>% 
  ggplot(aes(flux_fct, mean_area))+
  geom_point(aes(color = period), size = 2) + 
  geom_errorbar(aes(x = flux_fct, ymin = ci_lwr_area, ymax = ci_upr_area, color = period), size = 1) + 
  theme_bw()+ 
  ylab(expression(g~CO[2-eq]~yr^-1))+
  geom_hline(yintercept = 0, linetype ="dashed", "black")+
  scale_color_manual(values = c("#86BBD8", "#588157"))+
  scale_x_discrete(labels = c(expression(reservoir~CO[2-eq]), "burial", expression(river~CO[2]), expression(river~CH[4]), "NEP",  expression(soil~CH[4]), expression(tree~CH[4])))+
  theme(axis.text.x = element_text(vjust = 0, hjust  = 0.5,  size = 12, color = "black"), axis.text.y = element_text(size = 12, color = "black"), axis.title.x = element_blank(), axis.title.y = element_text(size =14), legend.title = element_text(size = 14), legend.text = element_text(size = 12))
dev.off()

png("../2-output-data/co2eq_draft3.png", width = 8, height = 5, units = "in", res = 300)
best_case_df %>% 
  filter(flux_unit == "CO2-eq", flux_fct %in% c("reservoir emissions", "river CH4 emissions", "river CO2 emissions")) %>% 
  ggplot(aes(flux_fct, mean_area))+
  geom_point(aes(color = period), size = 2) + 
  geom_errorbar(aes(x = flux_fct, ymin = ci_lwr_area, ymax = ci_upr_area, color = period), size = 1) + 
  theme_bw()+ 
  ylab(expression(g~CO[2-eq]~yr^-1))+
  geom_hline(yintercept = 0, linetype ="dashed", "black")+
  scale_color_manual(values = c("#86BBD8", "#588157"))+
  scale_x_discrete(labels = c(expression(reservoir~CO[2-eq]), expression(river~CO[2]), expression(river~CH[4])))+
  theme(axis.text.x = element_text(vjust = 0, hjust  = 0.5,  size = 12, color = "black"), axis.text.y = element_text(size = 12, color = "black"), axis.title.x = element_blank(), axis.title.y = element_text(size =14), legend.title = element_text(size = 14), legend.text = element_text(size = 12))
dev.off()

png("../2-output-data/co2eq_draft4.png", width = 10, height = 5, units = "in", res = 300)
best_case_df %>% 
  filter(flux_unit == "CO2-eq") %>% 
  ggplot(aes(flux_fct, mean_area))+
  geom_point(aes(color = period), size = 2) + 
  geom_errorbar(aes(x = flux_fct, ymin = ci_lwr_area, ymax = ci_upr_area, color = period), size = 1) + 
  theme_bw()+
  ylab(expression(g~CO[2-eq]~yr^-1))+
  geom_hline(yintercept = 0, linetype ="dashed", "black")+
  scale_color_manual(values = c("#86BBD8", "#588157", "#848484"))+
  scale_x_discrete(labels = c(expression(reservoir~CO[2-eq]), "burial", expression(exposed~sediment~CO[2]), expression(exposed~sediment~CH[4]), "drawdown ebullition", expression(eroded~sediment~CO[2]), expression(river~CO[2]), expression(river~CH[4]), "NEP", expression(soil~CH[4]), expression(tree~CH[4])))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust  = 1,  size = 12, color = "black"), axis.text.y = element_text(size = 12, color = "black"), axis.title.x = element_blank(), axis.title.y = element_text(size =14), legend.title = element_text(size = 14), legend.text = element_text(size = 12))
dev.off()

png("../2-output-data/co2eq_draft5.png", width = 10, height = 5, units = "in", res = 300)
worst_case_df %>% 
  filter(flux_unit == "CO2-eq") %>% 
  ggplot(aes(flux_fct, mean_area))+
  geom_point(aes(color = period), size = 2) + 
  geom_errorbar(aes(x = flux_fct, ymin = ci_lwr_area, ymax = ci_upr_area, color = period), size = 1) + 
  theme_bw()+
  ylab(expression(g~CO[2-eq]~yr^-1))+
  geom_hline(yintercept = 0, linetype ="dashed", "black")+
  scale_color_manual(values = c("#86BBD8", "#588157", "#848484"))+
  scale_x_discrete(labels = c(expression(reservoir~CO[2-eq]), "burial", expression(exposed~sediment~CO[2]), expression(exposed~sediment~CH[4]), "drawdown ebullition", expression(eroded~sediment~CO[2]), expression(river~CO[2]), expression(river~CH[4]), "NEP", expression(soil~CH[4]), expression(tree~CH[4])))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust  = 1,  size = 12, color = "black"), axis.text.y = element_text(size = 12, color = "black"), axis.title.x = element_blank(), axis.title.y = element_text(size =14), legend.title = element_text(size = 14), legend.text = element_text(size = 12))
dev.off()


png("../2-output-data/co2eq_draft6.png", width = 10, height = 5, units = "in", res = 300)
summary_df %>% 
  filter(flux == "exposed sediment CO2", flux_unit == "CO2-eq") %>% 
  mutate(variation = replace(variation, variation == "0.0833333333333333", "0.083") %>% fct_relevel(c("0.083", "1", "5", "10", "30", "100"))) %>% 
  ggplot(aes(variation, mean_area)) + 
  geom_point(color = "#848484", size = 2 )+
  geom_errorbar(aes(x = variation, ymin = ci_lwr_area, ymax = ci_upr_area), color = "#848484", size = 1) + 
  theme_bw()+
  ylab(expression(Exposed~sediment~CO[2]~emissions~g~CO[2-eq]))+
  xlab("Years to near zero emissions")+
  theme(axis.text.x = element_text(size = 12, color = "black"), axis.text.y = element_text(size = 12, color = "black"), axis.title = element_text(size =14), legend.title = element_text(size = 14), legend.text = element_text(size = 12))
dev.off()

png("../2-output-data/co2eq_draft7.png", width = 10, height = 7, units = "in", res = 300)
summary_df %>% 
  filter(flux_unit == "CO2-eq") %>% 
  mutate(variation = case_when(variation == "0.0833333333333333" ~ "0.083",
                               variation == "pore_0.0833333333333333" ~ "pore_0.083", 
                               variation == "no_pore_0.0833333333333333" ~"no_pore_0.083", 
                               TRUE ~ variation), flux_var = paste(flux, variation) %>% fct_relevel(c("reservoir burial NA", "reservoir emissions NA"))) %>% 
  ggplot(aes(flux_var, mean_area))+
  geom_point(aes(color = period), size = 2 )+
  scale_color_manual(values = c("#86BBD8", "#588157", "#848484"))+
  geom_errorbar(aes(x = flux_var, ymin = ci_lwr_area, ymax = ci_upr_area, color = period), size = 1) + 
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust  = 1,  size = 12, color = "black"), axis.text.y = element_text(size = 12, color = "black"), axis.title.x = element_blank(), axis.title.y = element_text(size =14), legend.title = element_text(size = 14), legend.text = element_text(size = 12))+
  ylab(expression(g~CO[2-eq]~yr^-1))+
  geom_hline(yintercept = 0, linetype ="dashed", "black")
dev.off()

unique(summary_df$flux)

png("../2-output-data/co2eq_draft8.png", width = 8, height = 5, units = "in", res = 300)
summary_df %>% 
  filter(flux_unit == "CO2-eq", flux %in% c("exposed sediment CO2", "exposed sediment CH4")) %>% 
  mutate(variation = case_when(variation == "0.0833333333333333" ~ "0.083",
                               variation == "pore_0.0833333333333333" ~ "pore_0.083", 
                               variation == "no_pore_0.0833333333333333" ~"no_pore_0.083", 
                               TRUE ~ variation) %>% fct_relevel(c("0.083", "1", "5", "10", "30", "100", "no_pore_0.083", "no_pore_1", "no_pore_5", "no_pore_10", "no_pore_30", "no_pore_100", "pore_0.083", "pore_1", "pore_5", "pore_10", "pore_30", "pore_100"))) %>% 
  ggplot(aes(variation, mean_area))+
  geom_point(aes(color = flux), size = 2 )+
  geom_errorbar(aes(x = variation, ymin = ci_lwr_area, ymax = ci_upr_area, color = flux), size = 1) + 
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust  = 1,  size = 12, color = "black"), axis.text.y = element_text(size = 12, color = "black"), axis.title.x = element_blank(), axis.title.y = element_text(size =14), legend.title = element_text(size = 14), legend.text = element_text(size = 12))+
  ylab(expression(g~CO[2-eq]~yr^-1))+
  geom_hline(yintercept = 0, linetype ="dashed", "black")
dev.off()

png("../2-output-data/co2eq_draft9.png", width = 10, height = 5, units = "in", res = 300)
middle_case_df %>% 
  filter(flux_unit == "CO2-eq") %>% 
  ggplot(aes(flux_fct, mean_area))+
  geom_point(aes(color = period), size = 2) + 
  geom_errorbar(aes(x = flux_fct, ymin = ci_lwr_area, ymax = ci_upr_area, color = period), size = 1) + 
  theme_bw()+
  ylab(expression(g~CO[2-eq]~yr^-1))+
  geom_hline(yintercept = 0, linetype ="dashed", "black")+
  scale_color_manual(values = c("#86BBD8", "#588157", "#848484"))+
  scale_x_discrete(labels = c(expression(reservoir~CO[2-eq]), "burial", expression(exposed~sediment~CO[2]), expression(exposed~sediment~CH[4]), "drawdown ebullition", expression(eroded~sediment~CO[2]), expression(river~CO[2]), expression(river~CH[4]), "NEP", expression(soil~CH[4]), expression(tree~CH[4])))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust  = 1,  size = 12, color = "black"), axis.text.y = element_text(size = 12, color = "black"), axis.title.x = element_blank(), axis.title.y = element_text(size =14), legend.title = element_text(size = 14), legend.text = element_text(size = 12))
dev.off()

middle_case_df %>% filter(period %in% c("before", "build"), flux_unit == "CO2-eq") %>% group_by(period) %>% summarize(sum_flux = sum(mean_area) %>% formatC(format = "e", digits = 2))

middle_case_df %>% filter(period %in% c("burp", "build"), flux_unit == "CO2-eq") %>% group_by(period) %>% summarize(sum_flux = sum(mean_area)) %>% ungroup() %>% pivot_wider(names_from = "period", values_from = "sum_flux") %>% mutate(burp/build)

middle_case_df %>% filter(period %in% c("burp", "build"), flux_unit == "CO2-eq") %>% group_by(period) %>% summarize(sum_flux = sum(ci_upr_area)) %>% ungroup() %>% pivot_wider(names_from = "period", values_from = "sum_flux") %>% mutate(burp/build)

middle_case_df %>% filter(period %in% c("before", "build"), flux_unit == "CO2-eq") %>% group_by(period) %>% summarize(sum_flux = sum(mean_area))
```

