---
title: "veazie"
author: "L. Naslund"
date: "2024-01-08"
output: html_document
---

The Veazie Dam was the further downstream of two dams removed on the Penobscot River in coastal central Maine.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rnaturalearth)
library(tidyverse)
library(sf)
library(raster)
library(mapview)
library(lme4)
library(MuMIn)
library(readxl)
library(merTools)
library(elevatr)

# required geospatial 
res <- st_read("../1-input-data/1-geospatial/reservoirs/veazie_res.shp")
ws <- st_read("../1-input-data/1-geospatial/watersheds/veaziewatershed.shp") %>% st_transform(st_crs(res))
riv <- st_read("../1-input-data/1-geospatial/post-removal/veazie_river.shp") %>% st_transform(st_crs(res))

mapview(res)+
  mapview(riv, col.regions = "red")

summary_df <- data.frame(period = character(), flux = character(), variation = character(), flux_unit = character(), annual = character(), area = character(), mean = double(), ci_lwr = double(), ci_upr = double())
```

```{r reservoir and catchment characteristics}
## g-res required inputs
# catchment area (km2)
sf_use_s2(FALSE)
ws_area <- st_area(ws)
print(paste0("Catchment area (km2): ", round(ws_area * 1e-6, 2)))

# gres required inputs
ws_pop <- 79672
print(paste0("Population in the catchment: ", ws_pop))

# community wastewater treatment
wwtp <- "secondary"
print(paste0("Wastewater treatment: ", wwtp))

# land cover
veazie_lc <- read.csv("../1-input-data/3-gres/veazie_lc.csv") %>% 
  mutate(gres_class = case_when(
    Legend == "Open Water" ~ "Water Bodies", 
    Legend == "Perennial Snow/Ice" ~ "Permanent Snow/Ice",
    Legend == "Developed, Open Space" ~ "Settlements", 
    Legend == "Developed, Low Intensity" ~ "Settlements", 
    Legend == "Developed, Medium Intensity" ~ "Settlements", 
    Legend == "Developed, High Intensity" ~ "Settlements", 
    Legend == "Barren Land" ~ "Bare Areas",
    Legend == "Deciduous Forest" ~ "Forest", 
    Legend == "Evergreen Forest" ~ "Forest", 
    Legend == "Mixed Forest" ~ "Forest", 
    Legend == "Shrub/Scrub" ~ "Grassland/Shrubland", 
    Legend == "Herbaceuous" ~ "Grassland/Shrubland", 
    Legend == "Hay/Pasture" ~ "Croplands", 
    Legend == "Woody Wetlands" ~ "Wetlands", 
    Legend == "Emergent Herbaceuous Wetlands" ~ "Wetlands",
  ))

print("Catchment landcover:")
lc <- veazie_lc %>% group_by(gres_class) %>% summarize(perc = sum(percent) %>% round(1)) 
print(lc)

# country
country <- "United States"
print(paste0("Country: ", country))

# coordinates
long <- -68.7011
lat <- 44.8322
print(paste0("Coordinates: ", lat, ",", long))

# reservoir area (km2)
res_area <- st_area(res) 
print(paste0("Reservoir Area (km2): ", res_area * 1e-6 ))

# mean depth (m)
mean_depth <- 4.42
print(paste0("Mean Depth (m): ", mean_depth))

# max depth (m)
max_depth <- 8.70
print(paste0("Max Depth (m): ", max_depth))

# littoral area (%) Output from g-res given from max and mean depth
lit_area <- 33.599
print(paste0("Littoral Area (%): ", lit_area))

# thermocline depth (m) Output from g-res
thermo_depth <- 0.1
print(paste0("Thermocline depth (m): ", thermo_depth))

# soil carbon
soil_carbon <- 5.11
print(paste0("Soil carbon under impounded area (kgC/m2): ", soil_carbon))

# reservoir wind
res_wind <- 5.31
print(paste0("Reservoir wind at 50 m: ", res_wind))

# water residence time (yrs) Output from g-res
wrt <- 0.0007
print(paste0("Water residence time (yrs): ", wrt))

# annual discharge
q_annual <- 345
print(paste0("Annual Discharge (m3/s): ", q_annual))

# p concentration (ug/L) Output from Cronan et al. 2012
p_conc <- 3
print(paste0("P Concentration (ug/L): ", p_conc))

# global horizontal radiance
rad <- 3.47
print(paste0("Global horizontal radiance: ", rad))

# mean temperature per month
temperature <- data.frame(Jan = -7.9, Feb = -6.7, Mar = -1.1, Apr = 5.4, May = 11.7, June = 17.0, July = 19.9, Aug = 18.8, Sept = 14.0, Oct = 8.4, Nov = 2.5, Dec = -4.8) %>% pivot_longer(cols = Jan:Dec, names_to = "Month", values_to = "Temperature")
print("Temperature per month (C)")
temperature
```

# Before

## Reservoir surface emissions

**G-res tool**

https://131.datatrium.com/fmi/webd/G-res%20Tool?script=ChoiceWebPage&param=Grestool&homeurl=https://g-res.hydropower.org

**Required data**

Must know

- reservoir age
- P concentration (ug/L) or population in the catchment and category of wastewater treatment
- water residence time or mean depth
- reservoir area (km^2)
- reservoir climate zone
- maximum depth
- mean depth or volume

Can be derived from other inputs

- % littoral area
- water residence time

Can be calculated from Google Earth Engine script given a delineated catchment with 
https://code.earthengine.google.com/db5542fcc97469961bfd890123aa55df

- mean annual air temperature (C)
- reservoir surface soil carbon content (kgC m^-2)
- cumulative global horizontal radiance (kWh m^-2 d-1)
- catchment area (km2)
- annual runoff (mm/yr)
- catchment landcover (%)
- k factor (soil erodibility factor)
- mean basin slope

Can delineate catchments in US reservoirs using the StreamStats tool 
https://streamstats.usgs.gov/ss/ 

```{r surface emissions}
# output from g-res: emissions in g CO2 eq by pathway
res_co2_diff_eq <- 3
res_ch4_diff_eq <- 79 * .89
res_ch4_eb_eq <- 79 * 0.11
res_ch4_degas_eq <- 0

# output from g-res: net GHG footprint and 95% CI
res_net_eq <- 64
res_net_2.5_eq <- 55
res_net_97.5_eq <- 75

res_net_interval <- max(c(res_net_97.5_eq-res_net_eq, res_net_eq - res_net_2.5_eq))

# calculate emissions and CI in g CO2 eq 
res_eq <- res_co2_diff_eq+res_ch4_diff_eq+res_ch4_eb_eq

res_eq_2.5 <- res_eq - res_net_interval
res_eq_97.5 <-  res_eq + res_net_interval

# calculate emissions and CI in g C 
res_g <- (res_co2_diff_eq * (12.01/44.01)) + ((res_ch4_diff_eq+res_ch4_eb_eq) *(12.01/16.04) * (1/34))

# to calculate CI in g C, I am assuming gases contribute to the CI in proportion to the CO2-eq contribution 
res_g_interval <- res_net_interval*(res_co2_diff_eq/res_eq)*(12.01/44.01) + res_net_interval*((res_ch4_diff_eq + res_ch4_degas_eq)/res_eq)*(1/34) * (12.01/16.04)

res_g_2.5 <- res_g - res_g_interval
res_g_97.5 <- res_g + res_g_interval

# summary data frame
res_emissions_summary <- data.frame(period = c("before"), flux = c("reservoir emissions"), variation = NA, flux_unit = c("CO2-eq", "C"), annual = "yes", area = "reservoir", mean = c(res_eq, res_g), ci_lwr = c(res_eq_2.5, res_g_2.5), ci_upr = c(res_eq_97.5, res_g_97.5))
```

## Reservoir carbon burial
```{r carbon burial, warnings = FALSE}
# sedimentation model
sed_df <- read.csv("../1-input-data/6-clow-data/clow_all.csv") %>% 
  mutate(sed_rate_m3_yr = sed_rate_mm_yr * 1e-3 * surface_area_m2, 
         log_sed_rate_m3_yr = log10(sed_rate_m3_yr), 
         log_surface_area_m2 = log10(surface_area_m2), 
         crop_perc = Crop_Perc/100, 
         forest_perc = Forest_Perc/100)

sed.mod <- lm(log_sed_rate_m3_yr ~ log_surface_area_m2 + slope_calc + forest_perc + crop_perc, data = sed_df)

veazie_sed_pred <- read.csv("../1-input-data/6-clow-data/veazie_sed_pred.csv") %>% 
  mutate(log_surface_area_m2 = log10(st_area(res) %>% as.numeric()), 
         crop_perc = lc %>% filter(gres_class == "Croplands") %>% pull(perc)/100, 
         forest_perc = lc %>% filter(gres_class == "Forest") %>% pull(perc)/100) %>% 
  dplyr::select(-Crop_Perc, -Forest_Perc, - dam) %>% 
  rename("slope_calc" = "slope")

predict(sed.mod, veazie_sed_pred, interval = "prediction")

veazie_sed_rate <- 10^(predict(sed.mod, veazie_sed_pred, interval = "prediction")[1])
veazie_sed_rate_sd <- (10^(predict(sed.mod, veazie_sed_pred, interval = "prediction")[1]) - 10^(predict(sed.mod, veazie_sed_pred, interval = "prediction")[2]))/2

# sediment c content model
c_content <- read.csv("../1-input-data/6-clow-data/c_content.csv") %>% 
  mutate(LAKE_ORIGIN_renamed = as.factor(LAKE_ORIGIN_renamed)) %>% 
  dplyr::select(sediment_C_pct, SOC_0_5, Wetlands90, log_Wetlands90, Barren30, KFactor, 
                log_WaterBody_Area_m2, log_Lake_Area_m2_EPA, LAKE_ORIGIN_renamed) %>% mutate(wetland_si = log10(Wetlands90 + 1))

c.mod <- lm(sediment_C_pct ~ SOC_0_5 + wetland_si + Barren30 + KFactor + log_Lake_Area_m2_EPA + LAKE_ORIGIN_renamed, data = c_content)


veazie_c_pred <- read.csv("../1-input-data/6-clow-data/veazie_soil.csv") %>% 
  rename("SOC_0_5" = "soc", "KFactor" = "kf") %>% 
  dplyr::select(-kw) %>% # using kf not kw
  mutate(wetland_si = log10(lc %>% filter(gres_class == "Wetlands") %>% pull(perc)/100 + 1), 
         Barren30 = lc %>% filter(gres_class == "Bare Areas") %>% pull(perc)/100, 
         log_Lake_Area_m2_EPA = log10(as.numeric(res_area)), 
         LAKE_ORIGIN_renamed = "RESERVOIR")

predict(c.mod, veazie_c_pred, interval = "prediction")

veazie_c_cont <- predict(c.mod, veazie_c_pred, interval = "prediction")[1] * 0.01
veazie_c_cont_sd <- ((predict(c.mod, veazie_c_pred, interval = "prediction")[1] - predict(c.mod, veazie_c_pred, interval = "prediction")[2])/2)*0.01

b = (veazie_c_cont_sd)^2/veazie_c_cont
a = (veazie_c_cont)^2/(veazie_c_cont_sd)^2
  
# dry bulk density
oc_dbd <- read.csv("../1-input-data/6-clow-data/oc_dbd.csv", header = F) %>% rename("OC" = "V1", "DBD" = "V2")

oc_dbd.mod <- lm(log(DBD)~log(OC), data = oc_dbd)

# burial efficiency
be <- read.csv("../1-input-data/6-clow-data/sed-rate_burial-efficiency.csv", header = FALSE) %>% rename("log_sed_rate_mm_yr" = "V1", "oc_be" = "V2")

be.mod <- lm(oc_be ~ log_sed_rate_mm_yr, data = be)

veazie_burial <- c()

for(i in 1:10000){
  print(i)
  
  # sample sedimentation rate and OC content
  temp_sed_rate <- rnorm(1, veazie_sed_rate, veazie_sed_rate_sd)
  temp_oc <- rgamma(1, shape = a, scale = b)
  
  # get dbd
  temp_oc_df <- data.frame(OC = temp_oc * 100)
  
  temp_dbd_distr <- exp(predict(oc_dbd.mod, temp_oc_df, interval = "prediction")[1]) * 1e6
  temp_dbd_distr_sd <- exp(predict(oc_dbd.mod, temp_oc_df, interval = "prediction")[1]) * 1e6 - exp(predict(oc_dbd.mod, temp_oc_df, interval = "prediction")[2]) * 1e6
  
  temp_dbd <- rnorm(1, temp_dbd_distr, temp_dbd_distr_sd)
  
  # get burial efficiency
  temp_be_df <- data.frame(log_sed_rate_mm_yr = log10(temp_sed_rate * (1/as.numeric(res_area)) * 1e3))
  
  temp_be_distr <- predict(be.mod, temp_be_df, interval = "prediction")[1] * 0.01
  temp_be_distr_sd <- ((predict(be.mod, temp_be_df, interval = "prediction")[1] * 0.01)-predict(be.mod, temp_be_df, interval = "prediction")[2] * 0.01)/2
  
  d = (temp_be_distr_sd)^2/temp_be_distr
  c = (temp_be_distr)^2/(temp_be_distr_sd)^2
  
  temp_be <- rgamma(1, shape = c, scale = d)
  
  temp <- temp_sed_rate * temp_dbd * temp_oc * temp_be * (1/as.numeric(res_area))  
  
  veazie_burial <- c(veazie_burial, temp)
}

veazie_burial <- sort(veazie_burial)

res_burial_summary <- data.frame(period = c("before"), flux = c("reservoir burial"), variation = NA, flux_unit = c("CO2-eq", "C"), annual = "yes", area = "reservoir", mean = c(-(veazie_burial[5000] * (44.01/12.01)), -veazie_burial[5000]), ci_lwr = c(-(veazie_burial[9750] * (44.01/12.01)), -veazie_burial[9750]), ci_upr = c(-(veazie_burial[250] * (44.01/12.01)), -veazie_burial[250]))
```

# Burp

## Exposed sediment CO2 emissions 
```{r exposed sed co2}
res_exp_sed_co2 <- read.csv("../2-output-data/exp_sed_co2.csv")

res_sed_co2_summary <- data.frame(period = c("burp"), flux = c("exposed sediment CO2"), variation = res_exp_sed_co2$y %>% as.character(), flux_unit = c(rep("CO2-eq", 6), rep("C", 6)), annual = "no", area = "exposed sediment area", mean = c(res_exp_sed_co2$mean_co2, res_exp_sed_co2$mean_c), ci_lwr = c(res_exp_sed_co2$ci_lower_co2, res_exp_sed_co2$ci_lower_c), ci_upr = c(res_exp_sed_co2$ci_upper_co2, res_exp_sed_co2$ci_upper_c))
```

## Exposed sediment CH4 emissions 
```{r exposed sed ch4}
thresh_0.01 <- read.csv("../2-output-data/exp_sed_ch4.csv")

pore_co2_eq <- 8102.2 * 1e-3 * (16.04/12.01)* 34
pore_c <- 8102.2 * 1e-3

res_sed_ch4_summary <- data.frame(period = c("burp"), flux = c("exposed sediment CH4"), variation = c(rep(paste0("no_pore_", thresh_0.01$y), 2), rep(paste0("pore_", thresh_0.01$y), 2)), flux_unit = rep(c(rep("CO2-eq", 6), rep("C", 6)), 2), annual = "no", area = "exposed sediment area", mean = c(thresh_0.01$mean_co2,  thresh_0.01$mean_c, thresh_0.01$mean_co2 + pore_co2_eq, thresh_0.01$mean_c + pore_c), ci_lwr = c(thresh_0.01$lwr_co2,  thresh_0.01$lwr_c, thresh_0.01$lwr_co2 + pore_co2_eq, thresh_0.01$lwr_c + pore_c), ci_upr = c(thresh_0.01$upr_co2, thresh_0.01$upr_c, thresh_0.01$upr_co2 + pore_co2_eq, thresh_0.01$upr_c + pore_c))
```

## Drawdown ebullition 
```{r drawdown eb}
res_ch4_eb_eq_2.5 <- res_ch4_eb_eq - max((res_net_eq - res_net_2.5_eq), (res_net_97.5_eq - res_net_eq))
res_ch4_eb_eq_97.5 <- res_ch4_eb_eq + max((res_net_eq - res_net_2.5_eq), (res_net_97.5_eq - res_net_eq))

drawdown_eb_prop <- c(0, 5, 10, 25, 100)

res_ch4_eb_g <- res_ch4_eb_eq * (1/34) * (12.01/16.04)
res_ch4_eb_g_2.5 <- res_ch4_eb_eq_2.5 * (1/34) * (12.01/16.04)
res_ch4_eb_g_97.5 <- res_ch4_eb_eq_97.5 * (1/34) * (12.01/16.04)

res_drawdown_eb_summary <- data.frame(period = c("burp"), flux = c("drawdown ebullition"), variation = drawdown_eb_prop %>% as.character(), flux_unit = c(rep("CO2-eq",5), rep("C", 5)), annual = "no", area = "reservoir", mean = c(res_ch4_eb_eq * drawdown_eb_prop, res_ch4_eb_g * drawdown_eb_prop) , ci_lwr = c(res_ch4_eb_eq_2.5 * drawdown_eb_prop, res_ch4_eb_g_2.5 * drawdown_eb_prop) , ci_upr = c(res_ch4_eb_eq_97.5 * drawdown_eb_prop, res_ch4_eb_g_97.5 * drawdown_eb_prop))
```

## Eroded sediment emissions 
```{r eroded sed}
# Collins et al. write that a combination of the CR Environmental Report and post-removal survey suggest that the combined volume of mobilized sediment from Veazie and Great Works dam was 15,000 m3 or smaller
# Fit a uniform distribution between 0 and 15,000 m3 

b = (veazie_c_cont_sd)^2/veazie_c_cont
a = (veazie_c_cont)^2/(veazie_c_cont_sd)^2

# assumming between 0 and 100% oxidized
min_ox_trans = 0 
max_ox_trans = 1

eroded_vec <- c()
for(i in 1:10000){
  print(i)
  
  # sample OC content
  temp_oc <- rgamma(1, shape = a, scale = b)
  
  # get dbd
  temp_oc_df <- data.frame(OC = temp_oc * 100)
  
  temp_dbd_distr <- exp(predict(oc_dbd.mod, temp_oc_df, interval = "prediction")[1]) * 1e6
  temp_dbd_distr_sd <- exp(predict(oc_dbd.mod, temp_oc_df, interval = "prediction")[1]) * 1e6 - exp(predict(oc_dbd.mod, temp_oc_df, interval = "prediction")[2]) * 1e6
  
  temp_dbd <- rnorm(1, temp_dbd_distr, temp_dbd_distr_sd)
  
  # parameters for g of sediment
  res_erode_g_max <- 15000 * temp_dbd
  res_erode_g_min <- 0
  
  temp <- (runif(1, min = res_erode_g_min, max = res_erode_g_max) * temp_oc * runif(1, min = min_ox_trans, max = max_ox_trans))/(res_area * 1e6)
  eroded_vec <- c(eroded_vec, temp)
}

eroded_vec <- sort(eroded_vec)

res_erode_summary <- data.frame(period = c("burp"), flux = c("eroded sediment emissions"), variation = NA, flux_unit = c("CO2-eq", "C"), annual = "no", area = "reservoir", mean = c(eroded_vec[5000] * (44.01/12.01), eroded_vec[5000]), ci_lwr = c(eroded_vec[250] * (44.01/12.01), eroded_vec[250]), ci_upr = c(eroded_vec[9750] * (44.01/12.01), eroded_vec[9750]))
```

# Build

## River surface CO2 emissions 
```{r river co2 emissions}
# Discharge is in the >100 m3/s bin
# Values are reported as g C m-2 d-

# Function to calculate the sum of squared differences between observed and target quantiles

gamma_params_diff <- function(params, quantiles_target) {
  shape <- params[1]
  scale <- params[2]
  
  quantiles_observed <- qgamma(c(0.25, 0.5, 0.75), shape = shape, scale = scale)
  
  sum((quantiles_observed - quantiles_target)^2)
}

# Target quantiles
quantiles_target <- c(1.5, 2.16, 4.78)

# Initial guess for shape and scale parameters
initial_params <- c(shape = 1, scale = 1)

# Optimize the parameters
optimized_params <- optim(par = initial_params, fn = gamma_params_diff, quantiles_target = quantiles_target)

# Extract optimized parameters
shape_optimized <- optimized_params$par[1]
scale_optimized <- optimized_params$par[2]

# Check the quantiles with the optimized parameters
quantiles_optimized <- qgamma(c(0.25, 0.5, 0.75), shape = shape_optimized, scale = scale_optimized)
cat("Optimized Quantiles:", quantiles_optimized, "\n")

x <- seq(0, 10, by = 0.1)
y <- dgamma(x, shape = shape_optimized, scale = scale_optimized)

plot(x, y)

riv_co2_vec <- sort(rgamma(10000, shape = shape_optimized, scale = scale_optimized)) *365

riv_co2_summary <- data.frame(period = c("build"), flux = c("river CO2 emissions"), variation = NA, flux_unit = c("CO2-eq", "C"), annual = "yes", area = "river", mean = c(riv_co2_vec[5000] * (44.01/12.01), riv_co2_vec[5000]), ci_lwr = c(riv_co2_vec[250] * (44.01/12.01), riv_co2_vec[250]), ci_upr = c(riv_co2_vec[9750] * (44.01/12.01), riv_co2_vec[9750]))
```

## River surface CH4 emissions 
```{r river ch4 emissions}
ch4_conc <- read.csv("../1-input-data/5-ch4-fluxes/GRiMe_concentrations_v2.csv") 
ch4_flux <- read.csv("../1-input-data/5-ch4-fluxes/GRiMe_fluxes_v2.csv")
ch4_flux_sites <- read.csv("../1-input-data/5-ch4-fluxes/GRiMe_sites_v2.csv")
ch4_source <- read.csv("../1-input-data/5-ch4-fluxes/GRiMe_sources_v2.csv")

usa <- ne_countries(country = "united states of america")

ch4_flux_sf <- ch4_flux %>% 
  left_join(ch4_flux_sites, by = c("Source_ID", "Site_ID", "Site_Name")) %>% 
  left_join(ch4_conc %>% 
              filter(FluxYesNo == "Yes") %>% 
              rename("Flux_Name" = "Conc_Name"), 
            by = c("Source_ID", "Site_ID", "Site_Name", "Flux_Name")) %>% 
  filter(is.na(Longitude)==FALSE, is.na(Channel_type)==TRUE) %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326) 

# filter sites in us
ch4_flux_sites_usa <- st_intersection(usa, ch4_flux_sf) %>% dplyr::select(Source_ID:geometry)
mapview(ch4_flux_sites_usa)

# Does it make sense to bin by discharge just because Hotckiss et al. 2015 did if discharge is not explanatory?
ch4_flux_sites_usa %>% ggplot(aes(Q, Diffusive_CH4_Flux_Mean))+geom_point()
ch4_flux_sites_usa %>% filter(Q < 500 & Diffusive_CH4_Flux_Mean<45) %>% ggplot(aes(Q, Diffusive_CH4_Flux_Mean))+geom_point()

ch4_flux_sites_usa <- ch4_flux_sites_usa %>% mutate(q_bin = case_when(
  (Q>=0 & Q<0.01) ~ "0-0.01",
  (Q>0.01 & Q<0.1) ~ "0.01-0.1",
  (Q>0.1 & Q<1) ~ "0.1-1",
  (Q>1 & Q<10) ~ "1-10",
  (Q>10 & Q<100) ~ "10-100",
  (Q > 100) ~ ">100"
)) 

# summary of diffusive CH4 by bin
ch4_flux_sites_usa %>% filter(is.na(Diffusive_CH4_Flux_Mean)==FALSE) %>% group_by(q_bin) %>% summarize(count = n())

# no measurements of ebullition in the Q range so assuming it is zero. This is another caveat we will need to address
ch4_flux_sites_usa %>% filter(is.na(Eb_CH4_Flux_Mean)==FALSE) %>% group_by(q_bin) %>% summarize(count = n())


ch4_flux_vec <- sort(sample(ch4_flux_sites_usa %>% filter(q_bin== ">100") %>% pull(Diffusive_CH4_Flux_Mean), 10000, replace = TRUE))
ch4_flux_vec_gCH4 <- ch4_flux_vec * 365 * 1e-3 * 16.04
ch4_flux_vec_gCO2eq <- ch4_flux_vec_gCH4 * 34

# CO2 flux
co2_vec <- sort(sample(ch4_flux_sites_usa %>% filter(is.na(CO2_Flux_Mean)==FALSE)  %>% filter(q_bin== "10-100") %>% pull(CO2_Flux_Mean), 10000, replace = TRUE))* 365 * 1e-3 * 44.01
co2_vec[250]
co2_vec[5000]
co2_vec[9750]

riv_ch4_summary <- data.frame(period = c("build"), flux = c("river CH4 emissions"), variation = NA, flux_unit = c("CO2-eq", "C"), annual = "yes", area = "river", mean = c(ch4_flux_vec_gCO2eq[5000], ch4_flux_vec_gCH4[5000] * (12.01 / 16.04)), ci_lwr = c(ch4_flux_vec_gCO2eq[250], ch4_flux_vec_gCH4[250] * (12.01 / 16.04)), ci_upr = c(ch4_flux_vec_gCO2eq[9750], ch4_flux_vec_gCH4[9750] * (12.01 / 16.04)))
```

## NEP 
```{r nee}
f_age <- function(a, b, e, age){
  return(e + a * (1-exp(b*age)))
}

df <- read.csv("../1-input-data/8-besnard-data/besnard_2018.csv") %>% mutate(age_func = f_age(263.2, -0.19, 0, Age)) %>% mutate(nutrient = as.factor(`NA.`))

# estimating NEP assuming that it is zero before the model predicts positive values
nep_mod <- lm(NEP ~ MAT + Age + age_func + nutrient, data = df)
summary(nep_mod)

nep_new <- data.frame(MAT = mean(temperature$Temperature), Age = seq(0, 100, by = 1), age_func = f_age(263.2, -0.19, 0, seq(0, 100, by = 1)), nutrient =as.factor("L"))
nep_pred <- predict(nep_mod, nep_new, interval = "prediction") %>% as.data.frame()

# becomes 0 a little before 5.7 years in so going to call it 6

nep_df <- data.frame(years = double(), mean_gC = double(), lwr_gC = double(), upr_gC = double())

year_vec <- c(25,50,100)
for(i in 1:length(year_vec)){
  yr <- year_vec[i]
  mean_gC <- integrate(function(x) as.data.frame(predict(nep_mod, newdata = data.frame(MAT = mean(temperature$Temperature), Age = x, age_func = f_age(263.2, -0.19, 0, x), nutrient =as.factor("L")), interval = "prediction", level = 0.95))$fit, lower = 6, upper = yr)$value
  lwr_gC <- -(integrate(function(x) -(as.data.frame(predict(nep_mod, newdata = data.frame(MAT = mean(temperature$Temperature), Age = x, age_func = f_age(263.2, -0.19, 0, x), nutrient =as.factor("L")), interval = "prediction", level = 0.95))$lwr), lower = 6, upper = yr))$value
  upr_gC <- integrate(function(x) as.data.frame(predict(nep_mod, newdata = data.frame(MAT = mean(temperature$Temperature), Age = x, age_func = f_age(263.2, -0.19, 0, x), nutrient =as.factor("L")), interval = "prediction", level = 0.95))$upr, lower = 6, upper = yr)$value
  
  temporary <- data.frame(years = yr, mean_gC = -mean_gC, lwr_gC = -upr_gC, upr_gC = -lwr_gC)
  
  nep_df <- nep_df %>% bind_rows(temporary)
}

# integrated carbon
nep_df <- nep_df %>% mutate(mean_gCO2 = mean_gC * (44.01/12.01), lwr_gCO2 = lwr_gC * (44.01/12.01), upr_gCO2 = upr_gC * (44.01/12.01))

# average annual rates
nep_df_rates <- nep_df %>% mutate(across(mean_gC:upr_gCO2, ~ .x / years))

nep_df_rates

res_nep_summary <- data.frame(period = c("build"), flux = c("NEP"), variation = nep_df_rates$years %>% as.character(), flux_unit = c(rep("CO2-eq", 3), rep("C", 3)), annual = "yes", area = "exposed sediment area", mean = c(nep_df_rates$mean_gCO2, nep_df_rates$mean_gC), ci_lwr = c(nep_df_rates$lwr_gCO2, nep_df_rates$lwr_gC), ci_upr = c(nep_df_rates$upr_gCO2, nep_df_rates$upr_gC))

res_nep_summary
```

## Soil CH4 emissions 
```{r soil ch4}
soil <- read.csv("../1-input-data/7-gattica-data/MethaneSoilForest.csv") %>% mutate(MAP_sq = MAP^2, MAT_sq = MAT^2) %>% filter(complete.cases(.))

soil_scale <- soil %>% mutate(across(c(MAP:soilpH, MAP_sq, MAT_sq), scale))

soil_mod <- lmer(Methane ~ MAP*MAT + soilpH + SOC + BulkDens + MAT_sq + SoilNit + MAP_sq + SoilSand + (1 + MAP*MAT + soilpH + SOC + BulkDens + MAT_sq + SoilNit + MAP_sq + SoilSand| Biome), data = soil_scale)

# I am not getting the same AICc
summary(soil_mod)
AICc(soil_mod, REML = FALSE)
AICc(soil_mod, REML = TRUE)

# need to get the predictors for sites
# MAP, MAT, soilpH, BulkDens, SoilNit, Soil Sand

veazie_soil <- read.csv("../1-input-data/7-gattica-data/veazie_soil.csv")

# going to have to manually scale                                                         
soil_data_new <- data.frame(MAP = (veazie_soil$veazie_map- mean(soil$MAP))/sd(soil$MAP),
                            MAT = (veazie_soil$veazie_mat - mean(soil$MAT))/sd(soil$MAT), 
                            soilpH = (veazie_soil$veazie_ph - mean(soil$soilpH))/sd(soil$soilpH), 
                            SOC = (veazie_soil$veazie_oc - mean(soil$SOC))/sd(soil$SOC), 
                            BulkDens = (veazie_soil$veazie_bd - mean(soil$BulkDens))/sd(soil$BulkDens), 
                            SoilNit = (veazie_soil$veazie_tn - mean(soil$SoilNit))/sd(soil$SoilNit), 
                            SoilSand = (veazie_soil$veazie_sand - mean(soil$SoilSand))/sd(soil$SoilSand), 
                            MAT_sq = (veazie_soil$veazie_mat^2 - mean(soil$MAT_sq))/sd(soil$MAT_sq), 
                            MAP_sq = (veazie_soil$veazie_map^2 - mean(soil$MAP_sq))/sd(soil$MAP_sq), 
                            Biome = "Temperate")


# returns values in kg Ch4 /ha/yr
soil_ch4_out <- predictInterval(soil_mod, soil_data_new, n.sims = 10000, level = 0.95) %>% as.data.frame()

soil_ch4_out <- soil_ch4_out %>% mutate(soil_c = fit * 0.1 * (12.01/16.04), 
                        soil_c_2.5 = lwr * 0.1 * (12.01/16.04), 
                        soil_c_97.5 = upr * 0.1 * (12.01/16.04), 
                        soil_co2_eq = fit * 0.1 * 34, 
                        soil_co2_eq_2.5 = lwr * 0.1 * 34, 
                        soil_co2_eq_97.5 = upr * 0.1 * 34)  
  
soil_ch4_out

res_soil_ch4_summary <- data.frame(period = c("build"), flux = c("soil CH4 emissions"), variation = NA, flux_unit = c("CO2-eq", "C"), annual = "yes", area = "exposed sediment area", mean = c(soil_ch4_out$soil_co2_eq, soil_ch4_out$soil_c), ci_lwr = c(soil_ch4_out$soil_co2_eq_2.5, soil_ch4_out$soil_c_2.5), ci_upr = c(soil_ch4_out$soil_co2_eq_97.5, soil_ch4_out$soil_c_97.5))
```

## Tree CH4 emissions 
```{r fvs inputs}
veazie_dem <- elevatr::get_elev_raster(res, z=14, expand = 2000)
veazie_masked <- terra::mask(veazie_dem, res)

elev <- mean(getValues(veazie_masked), na.rm = T) * 3.28084
elev

veazie_slope <- terrain(veazie_dem, opt = "slope")
veazie_slope_masked <- terra::mask(veazie_slope, res)

plot(veazie_slope_masked)

range(getValues(veazie_slope_masked), na.rm = T)
range(getValues(veazie_slope), na.rm = T)

slope <- mean(getValues(veazie_slope_masked), na.rm = T)
slope

veazie_aspect <- terrain(veazie_dem, opt = "aspect")
veazie_aspect_masked <- terra::mask(veazie_aspect, res)

aspect <- mean(getValues(veazie_aspect_masked), na.rm = T)
aspect

res %>% st_transform(crs = 4326) %>% st_centroid() %>% st_coordinates()
```

```{r tree ch4 angio, warnings = FALSE}
trees_k09_rm <- read_excel("../1-input-data/2-fvs/veazie-output-redmaple_k09.xlsx")
trees_k11_rm <- read_excel("../1-input-data/2-fvs/veazie-output-redmaple_k11.xlsx")
trees_k09_hem <- read_excel("../1-input-data/2-fvs/veazie-output-hemlock_k09.xlsx")
trees_k11_hem <- read_excel("../1-input-data/2-fvs/veazie-output-hemlock_k11.xlsx")

# confirmed that TPA gives you the right number of trees
# trees_k11_hem %>% group_by(Year) %>% summarize(trees = sum(TPA))

# DBH in inches Ht in ft convert to m2
sa_k09_rm <- trees_k09_rm %>% 
  mutate(surface_area = (DBH/12)*pi *EstHt *0.092903 * TPA, surface_area_3m = (DBH/12)*pi * 9.84 *0.092903 * TPA) %>% 
  group_by(StandID, Year) %>% summarize(total_surface_area = sum(surface_area), total_surface_area_3m = sum(surface_area_3m)) %>% 
  mutate(habitat = "wetland", tree_type = "angiosperm") %>% ungroup()

sa_k11_rm <- trees_k11_rm %>% 
  mutate(surface_area = (DBH/12)*pi *EstHt *0.092903 * TPA, surface_area_3m = (DBH/12)*pi * 9.84 *0.092903 * TPA) %>% 
  group_by(StandID, Year) %>% summarize(total_surface_area = sum(surface_area), total_surface_area_3m = sum(surface_area_3m)) %>% 
  mutate(habitat = "wetland", tree_type = "angiosperm") %>% ungroup()

sa_k09_hem <- trees_k09_hem %>% 
  mutate(surface_area = (DBH/12)*pi *EstHt *0.092903 *TPA, surface_area_3m = (DBH/12)*pi * 9.84 *0.092903 * TPA) %>% 
  group_by(StandID, Year) %>% summarize(total_surface_area = sum(surface_area), total_surface_area_3m = sum(surface_area_3m)) %>% 
  mutate(habitat = "wetland", tree_type = "gymnosperm") %>% ungroup()

sa_k11_hem <- trees_k11_hem %>% 
  mutate(surface_area = (DBH/12)*pi *EstHt *0.092903 *TPA, surface_area_3m = (DBH/12)*pi * 9.84 *0.092903 * TPA) %>% 
  group_by(StandID, Year) %>% summarize(total_surface_area = sum(surface_area), total_surface_area_3m = sum(surface_area_3m)) %>% 
  mutate(habitat = "wetland", tree_type = "gymnosperm") %>% ungroup()

# Pitz et al. 2018
# ug CH4/m2/hr -> g/m2/yr
angio_wet_tree_ch4 <- 567.9 * 8760 *1e-6
angio_wet_tree_ch4_sd <- 523.5 * 8760 *1e-6

# m2 in an acre
plot_area <- 4046.86

angio_ch4_df <- NULL

for(i in 1:10000){
  print(i)

  ch4_rate <- rnorm(1, angio_wet_tree_ch4, angio_wet_tree_ch4_sd)
  tree_sa <- sample(c(1, 2), size = 1, replace = TRUE)

  if(tree_sa == 1){
    temporary <- sa_k09_rm %>%
      mutate(ch4_all = total_surface_area * ch4_rate,
             ch4_3m = total_surface_area_3m * ch4_rate) %>%
      mutate(ch4_all_rate = ch4_all/plot_area, ch4_3m_rate = ch4_3m/plot_area) %>%
      summarize(mean_ch4_all = mean(ch4_all_rate),
        mean_ch4_3m = mean(ch4_3m_rate),
        ch4_all_rate_co2 = mean_ch4_all * 34,
        ch4_3m_rate_co2 = mean_ch4_3m * 34)
  }

  if(tree_sa == 2){
      temporary <- sa_k11_rm %>%
      mutate(ch4_all = total_surface_area * ch4_rate,
             ch4_3m = total_surface_area_3m * ch4_rate) %>%
      mutate(ch4_all_rate = ch4_all/plot_area, ch4_3m_rate = ch4_3m/plot_area) %>%
      summarize(mean_ch4_all = mean(ch4_all_rate),
        mean_ch4_3m = mean(ch4_3m_rate),
        ch4_all_rate_co2 = mean_ch4_all * 34,
        ch4_3m_rate_co2 = mean_ch4_3m * 34)
  }

  angio_ch4_df <- angio_ch4_df %>% bind_rows(temporary)
}


write.csv(angio_ch4_df, "../2-output-data/angio_ch4_emissions_3m.csv")
```

```{r tree ch4 gymno, warnings = FALSE}
# Machacova study of Scots pine is the only gymnosperm other than the bald cypress knees use the wet stem estimate
# use web plot digitizer to get IQR and then estimate SD using IQR/1.35. This is an underestimate of the SD because mean is higher than median

gymno_wet_tree_ch4 <- 0.166 * 8760 * 1e-6
gymno_wet_tree_ch4_sd <- ((0.225-0.042)/1.35)* 8760 * 1e-6

gymno_ch4_df <- NULL

for(i in 1:10000){
  print(i)
  ch4_rate <- rnorm(1, gymno_wet_tree_ch4, gymno_wet_tree_ch4_sd)
  tree_sa <- sample(c(1, 2), size = 1, replace = TRUE)

  if(tree_sa == 1){
       temporary <- sa_k09_hem %>%
      mutate(ch4_all = total_surface_area * ch4_rate,
             ch4_3m = total_surface_area_3m * ch4_rate) %>%
      mutate(ch4_all_rate = ch4_all/plot_area, ch4_3m_rate = ch4_3m/plot_area) %>%
      summarize(mean_ch4_all = mean(ch4_all_rate),
        mean_ch4_3m = mean(ch4_3m_rate),
        ch4_all_rate_co2 = mean_ch4_all * 34,
        ch4_3m_rate_co2 = mean_ch4_3m * 34)

  }

  if(tree_sa == 2){
       temporary <- sa_k11_hem %>%
      mutate(ch4_all = total_surface_area * ch4_rate,
             ch4_3m = total_surface_area_3m * ch4_rate) %>%
      mutate(ch4_all_rate = ch4_all/plot_area, ch4_3m_rate = ch4_3m/plot_area) %>%
      summarize(mean_ch4_all = mean(ch4_all_rate),
        mean_ch4_3m = mean(ch4_3m_rate),
        ch4_all_rate_co2 = mean_ch4_all * 34,
        ch4_3m_rate_co2 = mean_ch4_3m * 34)
  }

  gymno_ch4_df <- gymno_ch4_df %>% bind_rows(temporary)
}


write.csv(gymno_ch4_df, "../2-output-data/gymno_ch4_emissions_3m.csv")
```

```{r angio summary}
angio_ch4_df <- read.csv("../2-output-data/angio_ch4_emissions_3m.csv")

angio_ch4_all_mean <- (angio_ch4_df$mean_ch4_all %>% sort())[5000] * (12.01/16.04)
angio_ch4_all_2.5 <- (angio_ch4_df$mean_ch4_all %>% sort())[250] * (12.01/16.04)
angio_ch4_all_97.5 <- (angio_ch4_df$mean_ch4_all %>% sort())[9750] * (12.01/16.04)

angio_ch4_3m_mean <- (angio_ch4_df$mean_ch4_3m %>% sort())[5000] * (12.01/16.04)
angio_ch4_3m_2.5 <- (angio_ch4_df$mean_ch4_3m %>% sort())[250] * (12.01/16.04)
angio_ch4_3m_97.5 <- (angio_ch4_df$mean_ch4_3m %>% sort())[9750] * (12.01/16.04)

angio_ch4_all_mean_co2 <- (angio_ch4_df$ch4_all_rate_co2 %>% sort())[5000] * (12.01/16.04)
angio_ch4_all_2.5_co2 <- (angio_ch4_df$ch4_all_rate_co2 %>% sort())[250] * (12.01/16.04)
angio_ch4_all_97.5_co2 <- (angio_ch4_df$ch4_all_rate_co2 %>% sort())[9750] * (12.01/16.04)

angio_ch4_3m_mean_co2 <- (angio_ch4_df$ch4_3m_rate_co2 %>% sort())[5000] * (12.01/16.04)
angio_ch4_3m_2.5_co2 <- (angio_ch4_df$ch4_3m_rate_co2 %>% sort())[250] * (12.01/16.04)
angio_ch4_3m_97.5_co2 <- (angio_ch4_df$ch4_3m_rate_co2 %>% sort())[9750] * (12.01/16.04)
```

```{r gymno summary}
gymno_ch4_df <- read.csv("../2-output-data/gymno_ch4_emissions_3m.csv")

gymno_ch4_all_mean <- (gymno_ch4_df$mean_ch4_all %>% sort())[5000] * (12.01/16.04)
gymno_ch4_all_2.5 <- (gymno_ch4_df$mean_ch4_all %>% sort())[250] * (12.01/16.04)
gymno_ch4_all_97.5 <- (gymno_ch4_df$mean_ch4_all %>% sort())[9750] * (12.01/16.04)

gymno_ch4_3m_mean <- (gymno_ch4_df$mean_ch4_3m %>% sort())[5000] * (12.01/16.04)
gymno_ch4_3m_2.5 <- (gymno_ch4_df$mean_ch4_3m %>% sort())[250] * (12.01/16.04)
gymno_ch4_3m_97.5 <- (gymno_ch4_df$mean_ch4_3m %>% sort())[9750] * (12.01/16.04)

gymno_ch4_all_mean_co2 <- (gymno_ch4_df$ch4_all_rate_co2 %>% sort())[5000] * (12.01/16.04)
gymno_ch4_all_2.5_co2 <- (gymno_ch4_df$ch4_all_rate_co2 %>% sort())[250] * (12.01/16.04)
gymno_ch4_all_97.5_co2 <- (gymno_ch4_df$ch4_all_rate_co2 %>% sort())[9750] * (12.01/16.04)

gymno_ch4_3m_mean_co2 <- (gymno_ch4_df$ch4_3m_rate_co2 %>% sort())[5000] * (12.01/16.04)
gymno_ch4_3m_2.5_co2 <- (gymno_ch4_df$ch4_3m_rate_co2 %>% sort())[250] * (12.01/16.04)
gymno_ch4_3m_97.5_co2 <- (gymno_ch4_df$ch4_3m_rate_co2 %>% sort())[9750] * (12.01/16.04)
```

```{r tree ch4 summary}
res_tree_ch4_summary <- data.frame(period = c("build"),
                                   flux = c("tree CH4 emissions"),
                                   variation = c("offset by canopy angio", 
                                                 "offset by canopy angio", 
                                                 "whole trunk angio", 
                                                 "whole trunk angio", 
                                                 "three meters angio", 
                                                 "three meters angio", 
                                                 "offset by canopy gymno", 
                                                 "offset by canopy gymno", 
                                                 "whole trunk gymno", 
                                                 "whole trunk gymno", 
                                                 "three meters gymno", 
                                                 "three meters gymno"),
                                   flux_unit = rep(c("CO2-eq", "C"), 6),
                                   annual = "yes",
                                   area = "exposed sediment area",
                                   mean = c(0, 
                                            0, 
                                            angio_ch4_all_mean_co2, 
                                            angio_ch4_all_mean, 
                                            angio_ch4_3m_mean_co2, 
                                            angio_ch4_3m_mean,
                                            0, 
                                            0, 
                                            gymno_ch4_all_mean_co2, 
                                            gymno_ch4_all_mean, 
                                            gymno_ch4_3m_mean_co2, 
                                            gymno_ch4_3m_mean),
                                   ci_lwr = c(0, 
                                            0, 
                                            angio_ch4_all_2.5_co2, 
                                            angio_ch4_all_2.5, 
                                            angio_ch4_3m_2.5_co2, 
                                            angio_ch4_3m_2.5,
                                            0, 
                                            0, 
                                            gymno_ch4_all_2.5_co2, 
                                            gymno_ch4_all_2.5, 
                                            gymno_ch4_3m_2.5_co2, 
                                            gymno_ch4_3m_2.5),
                                   ci_upr = c(0, 
                                            0, 
                                            angio_ch4_all_97.5_co2, 
                                            angio_ch4_all_97.5, 
                                            angio_ch4_3m_97.5_co2, 
                                            angio_ch4_3m_97.5,
                                            0, 
                                            0, 
                                            gymno_ch4_all_97.5_co2, 
                                            gymno_ch4_all_97.5, 
                                            gymno_ch4_3m_97.5_co2, 
                                            gymno_ch4_3m_97.5))

res_tree_ch4_summary

```

# Summary
```{r create summary dfs}
res_area <- st_area(res) %>% as.numeric()
riv_area <- st_area(riv) %>% as.numeric()

area_df <- data.frame(area = c("reservoir", "exposed sediment area", "river"), 
                      area_m2 = c(res_area, (res_area - riv_area), riv_area))

# negative values are for uptake or burial positive values are for emissions

summary_df <- bind_rows(res_emissions_summary, res_burial_summary, res_sed_co2_summary, res_sed_ch4_summary, res_drawdown_eb_summary, res_erode_summary, riv_co2_summary, riv_ch4_summary, res_nep_summary, res_soil_ch4_summary, res_tree_ch4_summary) %>% left_join(area_df, by = "area") %>% mutate(mean_area = mean * area_m2, ci_lwr_area = ci_lwr * area_m2, ci_upr_area = ci_upr * area_m2) %>% mutate(flux_fct = fct_relevel(flux, c("reservoir emissions", "reservoir burial", "exposed sediment CO2", "exposed sediment CH4", "drawdown ebullition", "eroded sediment emissions", "river CO2 emissions", "river CH4 emissions", "NEP", "soil CH4 emissions", "tree CH4 emissions")))


write.csv(summary_df, "../2-output-data/veazie_summary_nep.csv", row.names = FALSE)
```

